Nup107 is a crucial regulator of torso-mediated metamorphic transition in Drosophila melanogaster

Version 1

Abstract

Abstract

Nuclear pore complexes (NPCs), composed of nucleoporins (Nups), affect nucleocytoplasmic transport, thus influencing cell division and gene regulation. Nup107 subcomplex members have been studied in housekeeping functions, diseases, and developmental disorders.

We report a unique regulatory function for Nup107 in metamorphic transition during Drosophila development. RNAi-mediated Nup107 depleted larvae were arrested in the third-instar larval stage and completely ceased to pupariate. The pupariation defect is primarily due to inhibited nuclear translocation and transcriptional activation by EcR. We unequivocally demonstrate the involvement of Nup107 in the transcription of the Halloween genes, modulating ecdysone biosynthesis and the EcR pathway activation. The regulation of EcR-mediated metamorphosis by the receptor tyrosine kinase, torso, is well documented. Accordingly, overexpression of the torso and MAP-kinase pathway activator, rasV12, in the Nup107 depletion background rescues the phenotypes, implying that Nup107 is an epistatic regulator of Torso-mediated activation of EcR signaling during metamorphosis.

Introduction

The multi-protein assembly of nucleoporins (Nups) constitutes nuclear pore complexes (NPCs). In eukaryotic cells, NPCs serve as molecular conduits for the trafficking of proteins and RNAs between the nucleus and the cytoplasm. The NPC composition was revealed from proteomic analyses conducted in yeasts and vertebrates, have contributed to our understanding, revealing that Approximately 30 distinct Nups are arranged in sub-structures and are distributed on different faces of NPCs (Cronshaw et al., 2002; Rout et al., 2000). The substructures are the outer cytoplasmic and inner nuclear rings, which surround a central inner ring called the scaffold ring, primarily aiding in the process of nucleocytoplasmic transport (Lin & Hoelz, 2019).

The largest Nup107 sub-complex, also called the Y-complex, is symmetrically located on both sides of the nuclear membrane. In metazoans, the Y-complex is composed of ten distinct Nups (ELYS, Nup160, Nup133, Nup107, Nup96, Nup85, Nup43, Nup37, Sec13, and Seh1). ELYS is a sub-stoichiometric Y-complex member and is present only on the nucleoplasmic side (D’Angelo & Hetzer, 2008; Morchoisne-Bolhy et al., 2015). Nup107, along with Nup133, forms the stalk of the Y-complex and is critically required for the Y-complex stability. The Nup107 complex plays a pivotal role in facilitating ELYS-coordinated post-mitotic NPC assembly (Boehmer et al., 2003; Walther et al., 2003) and messenger RNA (mRNA) export (Bai et al., 2004). Additionally, The Nup107 complex members actively participate in mitosis, contributing to the regulation of kinetochore-microtubule polymerization (Mishra et al., 2010; Zuccolo et al., 2007). The multitude of cellular processes where the Y-complex performs vital functions suggests it to be a central component in maintaining cellular homeostasis.

As a stable constituent of NPC, many Nups, including the members of the Y-complex, associate with chromatin and exert transcriptional regulation. Notably, interactions between active genes and Nups occurring predominantly within the nucleoplasm have been reported for dynamic Nups such as Nup98, ELYS, and Sec13 (Capelson, Liang, et al., 2010; Kalverda et al., 2010; Kuhn et al., 2019). In Drosophila, the dual nucleoporin, ELYS, governs the development, and ELYS RNAi-induced developmental defects are due to the reactivation of the Dorsal (NF-κB) pathway even during the late larval stages (Mehta et al., 2020).

Nup107 is associated with actively transcribing genes at the nuclear periphery (Gozalo et al., 2020). The disruption of Nup107 in zebrafish embryos leads to significant developmental anomalies, including the absence of the pharyngeal skeleton (Zheng et al., 2012). Moreover, the biallelic Nup107 mutations (D157Y and D831A) correlate well with clinical conditions such as microcephaly and steroid-resistant nephrotic Syndrome (Miyake et al., 2015). In Drosophila, Nup107 co-localizes with Lamin during meiotic division, and Nup107 depletion perturbs lamin localization, leading to a higher frequency of cytokinesis failure during male meiosis (Hayashi et al., 2016). Nup107 influences the regulation of cell fate in aged and transformed cells by modulating EGFR signaling and the nuclear trafficking of ERK protein (Kim et al., 2010).

Drosophila undergoes elaborate metamorphosis initiated by the neuropeptide prothoracicotropic hormone (PTTH) (Rewitz et al., 2013). Bilateral neurons projecting into the prothoracic gland (PG), when stimulated by the PTTH, induce Ecdysone production, which is subsequently released into the circulatory system for conversion by peripheral tissues into its active form, 20-hydroxyecdysone (20E) (Johnson et al., 2013; McBrayer et al., 2007; Shimell et al., 2018). The 20E binds to the ecdysone receptor, and the whole complex translocates into the nucleus and binds to chromatin to activate ecdysone inducible genes (Johnston et al., 2011; Kozlova & Thummel, 2002; Tennessen & Thummel, 2011). In the prothoracic gland, the primary neuroendocrine organ, PTTH signals through the receptor tyrosine kinase (RTK), Torso. The Torso-dependent activation of the MAP-kinase pathway is responsible for the production and release of ecdysone hormone. However, ecdysone synthesis can also be regulated by the EGFR pathway (Cruz et al., 2020; Yamanaka et al., 2013). While Nup107 modulates EGFR pathway activation, the involvement of EGFR and torso pathways in ecdysone-dependent metamorphosis is undeniable. We dissect the involvement of Nup107 in Torso-mediated signaling and underlying mechanisms during Drosophila metamorphosis.

In a reverse genetic RNAi screening for Nup107 complex members, we noted that Nup107 RNAi induces a significant developmental arrest at the third instar larval stage. Further analysis revealed that the Ecdysone Receptor (EcR) signaling pathway is perturbed, and EcR fails to translocate into the nucleus in Nup107 knockdown. The failure of the EcR nuclear localization upon Nup107 depletion is due to significantly reduced levels of the ecdysone hormone during the late third instar larval stage. Interestingly, overexpression of the torso and the rasV12 in Nup107 depleted larvae rescued the developmental arrest and subsequently initiated pupation. We propose that Nup107, an epistatic regulator of torso-pathway activation in the PG, enables ecdysone surge for efficient metamorphic transition.

Results

Nup107 is essential for larval to pupal metamorphic transition

Cell biological analyses in the mammalian cell culture system has shed light on critical regulatory roles of Nup107 in vertebrates. Yet its importance in development is poorly understood. In this context, we started the characterization of Drosophila Y-complex member Nup107 in greater detail. Utilizing RNAi lines (Nup107KK and Nup107GD), we performed ubiquitous depletion through Actin5C-GAL4. Interestingly, Nup107 depletion, leading to larvae arrest at the third instar stage, causing complete cessation of pupariation (120 h AEL, right panel, Figure 1A), was accompanied by an extension of larval feeding and growth periods. Quantitative assessment of Nup107 transcript levels in the Nup107GD and Nup107KK RNAi lines suggested efficient Nup107 knockdown (approximately 60-70%, Figure 1B). For further analyses, we generated polyclonal antibodies against the Nup107 amino-terminal antigenic fragment (amino acids 1–210; see methods for details). Purified anti-Nup107 polyclonal antibodies detected a band of approximately ∼100 kDa in lysates prepared from the control larval brain complex. The intensity of this ∼100 kDa band was significantly reduced in lysates prepared from organisms where Nup107 was knocked down ubiquitously (Figure 1C) using Actin5C-GAL4 driving Nup107 RNAi (denoted as ubiquitous hereon). Further, the immunostaining with Nup107 antibodies identified a conserved and robust nuclear rim staining pattern overlapping with mAb414 antibodies recognizing FG-nucleoporins (Figure S1A) and mRFP-tagged Nup107 expressed through its endogenous promoter (Figure S1B) in salivary gland tissues. These observations confirm the efficacy of Nup107 knockdown and provide a handle to assess levels and localization of Nup107 in affected tissues. To further investigate the role of Nup107 in development, we generated a Nup107 null mutant using CRISPR-Cas9-mediated gene editing. This comprehensive approach involving RNAi-mediated knockdown and CRISPR-Cas9 gene editing is expected to provide valuable insights into the significance of Nup107 in Drosophila development. The gRNAs targeting regions close to the start and stop codon of the nup107 gene generated knockout (Nup107KO) mutants, which were confirmed by sequencing and PCR (Figures S2A and S2B). However, the Nup107KO mutants could not be used as the Nup107KO homozygous shows lethality at the embryonic stage. So, we carried out most of the analyses hereon with Nup107 RNAi lines.

Nup107 depletion impairs metamorphosis.

(A) Growth profile of third instar larvae from Actin5C-Gal4 driven control and Nup107 knockdowns (Nup107GD and Nup107KK RNAi lines) at 96 h AEL (after egg laying) and 120 h AEL.

(B) Quantitation of Nup107 knockdown efficiency. Data are represented from at least three independent experiments. Statistical significance was derived from the Student’s t-test. Error bars represent SEM. **p = <0.001 and ****p = <0.0001.

(C) Immunodetection of Nup107 protein levels in third instar larval brain-complex lysates from control and Nup107 knockdown.

(D) Quantification of Nup107 protein levels seen in (C). Data are represented from at least three independent experiments. Statistical significance was derived from the Student’s t-test. Error bars represent SEM. ***p = <0.0002 and ****p = <0.0001.

(E) Comparison of pupariation profiles of control and Nup107 knockdown organisms.

Nup107 contributes significantly to ecdysone signaling

The depletion of Nup107 in Drosophila resulted in a distinct halt in growth and a developmental arrest at the third instar stage (Figure 1A). To discern the pathways affecting juvenile to adult developmental transition in Drosophila, we focused on levels of the sole insect steroid hormone, ecdysone.

We examined the localization of EcR in the late third instar salivary glands of both the control and Nup107-depleted larvae (using the Nup107KK line). Wild-type larvae exhibited normal EcR localization within the nucleus, but the nuclear translocation of EcR is perturbed in the Nup107-depleted larvae, with the bulk of the signal retained in the cytoplasm (Figures 2A and 2B). Quantitative analysis of EcR intensities in the cytoplasm and nucleus further established a significant decrease in EcR signals inside the nucleus when Nup107 is depleted (Figure 2C). This observation suggests that the Nup107 is required for EcR nuclear localization to mediate critical larval to pupal developmental transition. Furthermore, we noticed that significantly smaller size salivary glands and the brain complex in Nup107 depleted larvae (Figure S3).

Ubiquitous knockdown of Nup107 disrupts ecdysone signaling.

(A-B) Staining of third instar larval salivary glands from control (A) and ubiquitous Nup107 knockdown (B) with Ecdysone receptor (anti-EcR antibody, red) and Nup107 (anti-Nup107 antibody, green). DNA is stained with DAPI. Scale bars, 20 μm. Charts represent the line scan intensity profile of EcR (Red) and DAPI (Cyan) in the salivary gland nucleus region.

(C) Quantification of the nucleocytoplasmic ratio of EcR under control and Nup107 knockdown conditions. At least 45 nuclei were analyzed from 7 to 8 pairs of salivary glands. Statistical significance was derived from the Student’s t-test. Error bars represent SEM. ****p = < 0.0001.

(D-F) Analysis of ecdysone-inducible genes, EcR (D), Eip75A (E), and Eip74EF (F) expression, respectively, at the onset of metamorphosis (late third instar larvae stage).

Data are represented from at least three independent experiments. Statistical significance was derived from the Student’s t-test. The error bars represent the SEM. ***p = <0.0004 and ****p = <0.0001.

Prompted by the cytoplasmic accumulation of EcR, we investigated whether the mRNA levels of ecdysone-inducible genes were affected. When 20E binds with EcR, the activated EcR with enhanced affinity occupies the ecdysone response element (ERE) region in gene promoters. The ecdysone receptor is thought to function in a positive auto-regulatory loop, which may elevate EcR levels and sustain ecdysone signaling (Varghese & Cohen, 2007). Firstly, we examined the EcR transcript levels, revealing a reduction under Nup107-depletion conditions (Figure 2D). Subsequently, we measured the mRNA levels of known EcR target genes, Eip75A and Eip74EF, to find that the expression of each of these two target genes is reduced and correlates with Nup107 knockdown levels (Figures 2E and 2F). Similar results were observed in the depletion of Nup107 using the GD-based RNAi line (Figure S4). The defects observed during metamorphic transitions can be attributed to impaired ecdysone signaling upon Nup107-depletion, which prompted a closer examination of their regulatory relationship.

Nup107 is dispensable for the nuclear import of EcR in the target tissue

We established that the nuclear localization of EcR is impaired upon Nup107 knockdown (see Figure 2B). Typically, EcR nuclear localization follows an intricate mechanism where 20E binding to EcR allows its nuclear translocation and subsequent activation of its target genes (Cronauer et al., 2007; Johnston et al., 2011; Lenaerts et al., 2019). Biosynthesis of ecdysone hormone from dietary cholesterol requires a group of P450 enzymes coded by the Halloween genes, is a prerequisite for EcR activation (Kannangara et al., 2021; Niwa & Niwa, 2014). Moreover, the involvement of nucleoporin, Nup358, in facilitating nuclear transport of Met juvenile hormone receptors is well documented (He et al., 2017). Consequently, we posited the hypothesis that Nup107 either disallows active 20E-EcR complex formation by affecting 20E biosynthesis in ubiquitous Nup107 knockdown scenarios or directly regulates EcR nuclear translocation in the target tissue.

We chose salivary glands to address these hypotheses and depleted Nup107 using salivary glands-specific AB1-GAL4 and PG-specific Phm-GAL4. Surprisingly, in contrast to the ubiquitous knockdown of Nup107, nuclear localization of EcR remained unaffected in salivary gland-specific Nup107 knockdown (Figures 3A, 3B, S5A and S5B). Interestingly, the PG-specific Nup107 knockdown phenocopied ubiquitous Nup107 knockdown induced EcR nuclear localization defects (Figures 3C and S5C).

Nup107 regulates Ecdysone receptor-dependent signaling.

(A-C) Detection of and quantitation of nucleocytoplasmic distribution of EcR (anti-EcR antibody, red) and Nup107 (anti-Nup107 antibody, green) in control (A), salivary gland-specific Nup107 depletion (B), and prothoracic gland-specific Nup107 depletion (C) from third instar larval salivary gland nuclei. DNA is stained with DAPI. Scale bars, 20 μm. Charts represent the line scan intensity profile of EcR (Red) and DAPI (Cyan) in the salivary gland nucleus region.

(D) EcR nucleo-cytoplasmic quantification ratio from the salivary gland and prothoracic gland-specific Nup107 knockdown, respectively. At least 45 nuclei were analyzed from 7 to 8 pairs of salivary glands. Statistical significance was derived from the Student’s t-test. Error bars represent SEM. ****p = <0.0001, and ns is non-significant.

(E-F) Quantitation of expression of Eip75A (E) and Eip74EF (F) ecdysone-inducible genes at the onset of metamorphosis (RNA isolated from late third instar larvae of control and prothoracic gland-specific Nup107 depletion). Data are represented from at least three independent experiments. Statistical significance was derived from the Student’s t-test. Error bars represent SEM. **p = <0.008 and ****p = <0.0001.

Quantification of EcR signals from Nup107 depleted late third instar salivary gland cells and comparison with control salivary glands indicates that the nuclear/cytoplasmic ratios are drastically reduced in case of PG-specific depletion but unaltered in salivary gland-specific Nup107 depletion (Figures 3D and S5D). Consistent with these observations, expression levels of ecdysone-inducible genes Eip75A and Eip74EF were significantly reduced in PG-specific Nup107 knockdown (Figures 3E, 3F, S5E and S5F). In accordance with unperturbed EcR nuclear translocation, salivary glands-specific depletion of Nup107 yielded no discernible differences in larval growth and pupariation compared to the control (Figures S6A and S6C). However, the PG-specific knockdown induced an extended third instar stage lifespan (10-12 days after egg laying, Figures S6B and S6C). The observed decrease in ecdysone-inducible gene expression during late third-instar developmental stages can explain the potential impairment of metamorphosis induction seen upon ubiquitous or PG-specific Nup107 knockdown. In addition to the reduced size of the salivary gland and brain complex, we also noticed a compromise in the size of the Prothoracic gland upon Nup107 knockdown (Figure S7).

These observations suggest that Nup107 exerts a regulation on ecdysone biosynthesis and active 20E-EcR complex formation rather than playing a direct role in EcR nuclear translocation.

Nup107 exerts control on the EcR pathway through Ecdysone level regulation

We reasoned that Nup107 may regulate the ecdysone biosynthesis in PG to induce the larval stage growth arrest. We delved into analyzing the effect of Nup107 knockdown on ecdysone production. The considerable decrease in PG size due to Nup107 knockdown (Figure S7) can potentially reduce 20E production. This prompted us to explore the potential role of Nup107 in influencing ecdysone production, and we assessed the impact of Nup107 knockdown on ecdysone biosynthesis.

Utilizing an ELISA-based detection method, we assessed 20E levels in larvae at 96 hours and 120 hours after egg laying (AEL). Larvae from different experimental conditions, including control, ubiquitous Nup107-depletion, and PG-specific Nup107-depletion, were used in this analysis. Strikingly, the results indicated a substantial decrease (∼ 3-fold with ubiquitous and ∼10-fold for PG-specific knockdown) in total 20E levels upon Nup107 knockdown, particularly at 120 AEL which coincides with the 20E surge seen during metamorphosis (Figures 4A and S8A). This observation is crucial and suggests a potential defect in ecdysone biosynthesis in Nup107-depleted organisms.

Nup107 critically regulates the expression of ecdysone-biosynthetic genes.

(A) ELISA measurements of whole-body 20-hydroxyecdysone (20E) levels in control, ubiquitous (Actin5C-Gal4), and prothoracic gland-specific (Phm-Gal4) Nup107 depletion at 96 and 120 h AEL. Data are represented from at least three independent experiments. Statistical significance was derived from the Student’s t-test. Error bars represent SEM. **p = <0.0013, ***p = < 0.0009 and ns is non-significant.

(B) Schematic representation of a Prothoracic gland cell showing genes involved in ecdysone biosynthesis from cholesterol.

(C-G) Quantification of ecdysone-biosynthetic gene expression levels of Spookier (C), Phantom (D), Disembodied (E), Shadow (F), and Shade (G) in cDNA isolated from control, and Nup107 knockdown late third instar larvae. Data are represented from at least three independent experiments. Statistical significance was derived from the Student’s t-test. Error bars represent SEM. **p = <0.001, ***p = <0.0005 and, ****p = <0.0001.

As depicted in Figure 4B, the PTTH hormone signaling in the PG upregulates the expression of Halloween genes (spookier, phantom, disembodied, and shadow) responsible for ecdysone biosynthesis, and the shade gene product is required for active 20E generation in peripheral tissues (Christensen et al., 2020; McBrayer et al., 2007; Shimell et al., 2018). We analyzed the Halloween genes transcript level in Nup107-depleted larvae and observed a significant downregulation for each of the Halloween quartet genes mentioned earlier. (Figures 4C-4F and S8B-S8E). Further, the level of the shade gene required for converting precursor ecdysone to active 20E was also reduced twofold in Nup107-depleted larvae (Figures 4G and S8F).

20E rescues Nup107-dependent EcR localization defects

The observed correlation between the expression levels of ecdysone biosynthetic genes and reduced levels of 20E upon Nup107 knockdown strongly suggests a role for Nup107 in 20E biosynthesis, active 20E-EcR complex formation, and its nuclear translocation. It is not surprising in this context that exogenous ecdysone supplementation through larval food rescues pupariation blocks arising from ecdysone deficiency (Garen et al., 1977; Ou et al., 2016; Shimell et al., 2018). We experimented the same under Nup107-depletion conditions and supplemented exogenous ecdysone to PG-specific Nup107-depleted larvae by feeding them with a diet enriched with 20-hydroxyecdysone (0.2 mg/ml). Supplementation of 20E to Nup107-depleted larvae significantly (70-80%) alleviated the developmental arrest, and the onset of pupariation was comparable to the control (Table S1). However, none of the pupae could eventually eclose successfully, probably due to other effects of Nup107 depletion. We asked if nuclear translocation of EcR can also be rescued by exogenous supplementation of 20E. While incubation of salivary glands of late third instar larva in S2 media control alone did not rescue EcR localization in any of the Nup107-knockdown genotypes (Figures 5A-5C and S9A), we noticed a complete EcR nuclear translocation rescue in ubiquitous as well as PG-specific Nup107-depleted salivary glands when incubated with 20E (Figures 5D-5F and S9B).

20E supplementation rescues Nup107 depletion-specific EcR signaling defects.

(A-F) Visualization of the nucleocytoplasmic distribution of EcR (anti-EcR antibody, red) without 20E (A-C) and with 20E (D-F) treatment in larval salivary glands of control, ubiquitous (Actin5C-Gal4) Nup107 knockdown, and prothoracic gland-specific (Phm-Gal4) Nup107 knockdown. DNA is stained with DAPI. Scale bars, 20 μm. Charts represent the line scan intensity profile of EcR (Red) and DAPI (Cyan) in the salivary gland nucleus region.

(G-H) Comparative quantification of expression ecdysone-inducible genes Eip75A (G) and Eip74EF (H) from 20E treated salivary glands. Data are represented from at least three independent experiments. Statistical significance was derived from the Student’s t-test. Error bars represent SEM. *p = <0.02, **p = <0.002, ***p = <0.0008 and ns is non-significant.

The exogenous supplementation of 20E leads to active 20E-EcR complex formation as mRNA levels of the Eip75A and Eip74EF target genes were rescued significantly back to normal levels (Figures 5G and 5H). Overall, these findings highlight the important regulatory function of Nup107 in the ecdysone signaling pathway.

Torso is an effector of Nup107 mediated functions in metamorphosis

Next, we explored the mechanism of Nup107-driven regulation of 20E levels and metamorphosis. Receptor tyrosine kinase family of cell surface receptors signal by binding to ligands (growth factors and hormones) and participate in metabolism, cell growth, and development. Among these RTKs, the Torso, belonging to the platelet-derived growth factor receptor (PDGFR) class, plays a significant role during metamorphosis by serving as a receptor for the neuropeptide PTTH in the Drosophila brain (Sopko & Perrimon, 2013).

Functional engagement of PTTH-Torso activates the MAP kinase pathway, involving components of Ras, Raf, MEK, and extracellular signal-regulated kinase (ERK), thereby initiating metamorphosis (Figure 6A). The torso knockdown in the prothoracic glands resulted in the developmental arrest phenotype synonymous with what is observed with Nup107 knockdown. Feeding 20E to torso-depleted larvae completely rescued developmental delay and normal growth phenotypes (Rewitz et al., 2009). We observed a significant decrease in the torso levels (∼ four-fold) when Nup107 was depleted ubiquitously using Actin5C-GAL4 (Figure 6B), suggesting an epistatic regulation by Nup107 on the torso. Further, the phenotypic similarity between Nup107 and torso-depletion scenarios and diminished torso level upon Nup107-depletion prompted us to investigate whether the torso is an effector of Nup107.

Torso and Nup107 act synergistically to activate the ecdysone signaling.

(A) A model of the Torso pathway and its components.

(B) Quantitation of torso transcript levels from control and Nup107 depleted larvae (ubiquitous depletion using Actin5C-Gal4). Data are represented from at least three independent experiments. Statistical significance was derived from the Student’s t-test. Error bars represent SEM. ****p = <0.0001.

(C-D) Comparison of pupariation profiles of control, Nup107 knockdown, and torso and rasV12 overexpressing rescue organisms.

(E-G) Detection and quantitation of nucleocytoplasmic distribution of EcR (anti-EcR antibody, red) and Nup107 (anti-Nup107 antibody, green) in control, torso overexpressing ubiquitous Nup107 knockdown (Actin5C-Gal4>Nup107KK; UAS-torso) and torso overexpressing PG-specific Nup107 knockdown (Phm-Gal4>Nup107KK; UAS-torso) third instar larval salivary gland nuclei. DNA is stained with DAPI. Scale bars, 20 μm. Charts represent the line scan intensity profile of EcR (Red) and DAPI (Cyan) in the salivary gland nucleus region.

(H-I) Quantification of expression of Eip75A (H) and Eip74EF (I) ecdysone-inducible genes, respectively. Data are represented from at least three independent experiments. Statistical significance was derived from the Student’s t-test. Error bars represent SEM. *p = <0.03, **p = <0.008, ****p = <0.0001 and ns is non-significant.

Overexpression of the torso and or rasV12 have been utilized to rescue torso pathway-mediated defects (Cruz et al., 2020; Rewitz et al., 2009). We probed the possibility of Nup107 phenotype rescue by ubiquitous or PG-specific overexpression of the torso and rasV12. Overexpression of either torso or rasV12 in Nup107-depletion background completely rescued the pupariation defects (Figures 6C and 6D). The torso overexpression-mediated rescue of Nup107-depletion phenotypes is specific as overexpression of Egfr (another RTK implicated in metamorphosis) and Usp (co-receptor with EcR) in Nup107 depletion background could not rescue the pupariation defects (data not shown). The remarkable rescue demonstrated by torso overexpression prompted us to analyze the status of EcR nuclear translocation, ecdysone biosynthesis, and ecdysone-inducible gene expression. The restoration of EcR nuclear translocation (Figures 6E-6G), the ecdysone biosynthesis genes, spookier, phantom, disembodied, and shadow (Figures S10B-S10E), and ecdysone target genes Eip75A and Eip74EF (Figures 6H and 6I) to control levels indicated that torso could efficiently rescue Nup107 phenotypes.

These observations suggest that the torso is a downstream effector of Nup107 functions, and torso-dependent signaling, responsible for 20E synthesis in PG and metamorphosis, is regulated by Nup107 levels.

Discussion

Apart from its importance in mRNA export and cell division at the cellular level, the NPCs Y-complex member Nup107, when mutated, correlates with developmental abnormalities, microcephaly, and nephrotic syndrome (Miyake et al., 2015; Zheng et al., 2012). In several human cancer types, Nup107 exhibits high expression and serves as a biomarker for hepatocellular carcinoma. A strong structural and functional conservation between human and Drosophila Nup107 proteins allows Drosophila to be used to model to gain mechanistic insights in Nup107 functions in human diseases (Shore et al., 2022; Weinberg-Shukron et al., 2015). We observe that Nup107 is involved in critical developmental transitions from larva to pupa during Drosophila development as pupariation is completely arrested (Figure 1). The foraging third instar larva must acquire a critical weight and sufficient energy stores before it can metamorphose into a non-feeding pupa, molting into a healthy adult. The activation of the Torso receptor and sequent surge in ecdysone synthesis is essential for larval to pupal transition (Christensen et al., 2020; Hao et al., 2021; Luo et al., 2024). The elevated ecdysone levels trigger the translocation of the heterodimeric ecdysone receptor [in complex with Ultraspiracle (Usp)] into the nucleus to induce pupariation-specific transcriptional programming (Johnston et al., 2011). Similarly, when Nup107 is depleted, nuclear translocation of EcR is abolished, arresting metamorphosis at the level of the third instar larval stage (Figure 2).

Translocation of nuclear receptor proteins like the ecdysone receptor to the nucleus is a crucial step for transcriptional activation. Nup358, present at the cytoplasmic filament face of NPC, plays a vital role in the nuclear import of Met (juvenile hormone receptor) along with Hsp83 (He et al., 2017) to help maintain the larval stages of development. The marked absence of EcR from the nucleus was intriguing as Nup107 generally participates in the nuclear export process. Perhaps Nup107 regulates pupariation by modulating ecdysone synthesis, ecdysone signaling, and nuclear translocation of EcR. Surprisingly, Nup107 is dispensable for the nuclear translocation of EcR in the target tissue, the salivary gland cells (Figure 3). Nup107 exerts strong control over the expression of Halloween genes involved in ecdysone biosynthesis, resulting in diminished 20E titer, poor EcR activation, and delayed larva-to-pupa transition (Figures 4 and S6). The effect of Nup107 on Halloween genes adheres well with pupariation defects. These observations are in concurrence with reports of low ecdysone levels disrupting the pupariation process, leading to a halt in insect development (Christensen et al., 2020; Cruz et al., 2020).

In addition to ecdysone signaling, insect metamorphosis also has a strong contribution from receptor tyrosine kinase RTK pathway signaling. Notably, Torso signaling is essential for embryonic development and metamorphosis in Drosophila. The receptor expression level impacts the signaling output: high receptor levels trigger a robust and transient signal, while lower levels result in a weaker, sustained signal (Jenni et al., 2015; Konogami et al., 2016). Accordingly, the reduced ecdysone levels in the torso knockdown prothoracic glands induced pupariation delays (Rewitz et al., 2009). We observe a similar defect with Nup107 knockdown, indicating a possible reduction in torso levels. Interestingly, we observed significantly reduced torso levels in Nup107-depleted organisms, suggesting an essential upstream function for Nup107 in regulating RTK pathway activation and the associated Drosophila metamorphosis process. In the contrasting observation, Nup107 depleted larval tissues (salivary glands, brain complex, and prothoracic gland) are significantly smaller in size (Figures S3 and S7). It is important to know how Nup107 contributes to organ size maintenance and if a crosstalk with RTK signaling is required in this context.

The developmental delays observed in Nup107 knockdown larvae can be restored by exogenous supplementation of 20E, suggesting that Nup107 can modulate directly or indirectly the Torso receptor activity for ecdysone production and developmental regulation. Accordingly, we successfully rescued developmental delays by ubiquitous and PG-specific torso overexpression. Perhaps overexpression of the torso pathway mediator molecules can rescue the Nup107 developmental delay phenotypes. The oncogenic variant of the Ras-GTPase (rasV12) has been explored in a similar analysis with torso pathway mutants (Cruz et al., 2020; Rewitz et al., 2009). The alleviation of the developmental delays in the Nup107 knockdown background upon rasV12 overexpression is an indication of Nup107 serving as an epistatic regulator of the torso pathway in metamorphic transitions (Figure 6).

In essence, our findings indicate that Nup107 influences pupariation timing by regulating the torso levels, it’s signaling, and ecdysone biosynthesis (Figure 7). Previous research has shown that nuclear pore complexes (NPCs) are essential for maintaining global genome organization and regulating gene expression (Capelson, Doucet, et al., 2010; Capelson, Liang, et al., 2010; Iglesias et al., 2020). Particularly, Nup107 interacts with chromatin and targets active gene domains to regulate gene expression (Gozalo et al., 2020). Thus, Nup107 exerting its effects on torso transcription is the primary regulatory event in the Drosophila metamorphosis. It is important to note that the synthesis and availability of Torso ligand, PTTH, may not be affected by the Nup107 since torso or downstream effector rasV12 overexpression is sufficient to rescue the developmental arrest phenotype. It is thus crucial to further delineate the mechanism of Nup107-dependent regulation on torso pathway activation. The whole genome transcriptomics from PG can help shed more light on the regulatory roles of Nup107. This informationwill offer valuable insights into how nucleoporins regulate gene expression and serve as a model for elucidating how they govern the temporal specificity of developmental processes in organisms. These analyses will help establish a link between Nup107, the PTTH-PG axis, and the regulation of maturation timing. Our observations indicate critical roles for Nup107 in both torso-ecdysone interplay in metamorphosis and torso-independent mechanism for organ size maintenance. Together they add valuable information to hitherto unknown functions of the Nup107 in organismal development. Further investigations are required to identify the interactors of Nup107 involved in these coordination mechanisms in developmental transition.

Theoretical Model of Nup107 functions in metamorphosis:

During metamorphosis, the prothoracic gland (PG) responds to prothoracicotropic hormone (PTTH) via Torso receptors. The MAP kinase pathway involving Raf, MEK, and ERK, initiated by Ras, leads to Halloween gene expression responsible for Ecdysone synthesis and release. Ecdysone is converted to active 20-hydroxyecdysone (20E) in peripheral tissues. The binding of 20E to the Ecdysone receptor (EcR) allows EcR nuclear translocation and EcR pathway activation, culminating in target gene expression facilitating the metamorphic transition. Nup107 depletion negatively impacts Torso levels and Torso pathway activation, inducing pupariation arrest, which can be rescued by autonomous activation of the Torso pathway. Image created with BioRender.com/x18z538.

Materials and methods

Key resources table

Fly stocks and genetics

Experimental Drosophila melanogaster stocks were reared on a standard cornmeal diet (Nutri-Fly Bloomington formulation) under controlled conditions of 25°C and 60% relative humidity unless otherwise specified. Fly lines used in this study were sourced from the Bloomington Drosophila Stock Centre (BDSC) at Indiana University or from the Vienna Drosophila Resource Center (VDRC), which are listed in the Key resources table. The UAS-GFP lines were received as a gift from Dr. Varun Chaudhary (IISER Bhopal, India). Control groups were generated through crosses between the driver line and w1118 flies. For RNA interference (RNAi) experiments, crosses were maintained at 29°C to optimize GAL4 expression. The Nup107GD line, in conjunction with Actin5C-GAL4, was specifically cultivated at 23°C to generate third-instar larvae for experimental purposes.

Transgenic fly generation

In the generation of transgenic flies containing gRNA, we employed a systematic approach. The online tool available at http://targetfinder.flycrispr.neuro.brown.edu was used for the initial sgRNA design, ensuring zero predicted off-targets. Subsequently, we utilized an additional tool available at https://www.flyrnai.org/evaluateCrispr/ to evaluate and score the predicted efficiency of sgRNAs in the targeted region. The most efficient sgRNAs, demonstrating high specificity with no predicted off-targets, were selected and cloned into the pCFD4 vector. The Fly Facility services at the Centre for Cellular and Molecular Platforms, National Center for Biological Sciences (C-CAMP-NCBS), Bangalore, India, were utilized to clone and generate transgenic flies. Primer sequences employed in this study can be found in the Key resources table.

CRISPR –Cas9 mediated mutant Generation

Virgin nanos.Cas9 (BL-54591) flies were crossed with males of gRNA transgenic lines, and approximately 10 F1 progeny males were crossed with balancer flies specific for the gene of interest. A single-line cross was established using the F2 progeny to assess the deletion of nup107. Subsequently, F3 progeny were subjected to nested PCR to confirm deletions. Positive individuals were further tested for the presence of gRNA and Cas9 transgenes, and flies lacking both were propagated into stocks.

Genomic DNA Isolation

Approximately ten flies were homogenized in 250 µL of solution A, comprising 0.1 M Tris-HCl, pH 9.0, 0.1 M EDTA, and 1% SDS, supplemented with Proteinase-K. The homogenate was incubated at 70°C for 30 min. Subsequently, 35 µL of 8 M potassium acetate was added, mixed gently without vortexing, and incubated on ice for 30 min. The homogenate was then centrifuged at 13,000 rpm at 4°C for 15 min, and the supernatant was carefully collected without disturbing precipitates or the interphase. 150 µL of Isopropanol was added to the supernatant and incubated on ice for 15 min. After centrifugation at 13,000 rpm for 5 min, the supernatant was discarded, and the pellet was washed with 1 mL of 70% ethanol by centrifuging at 13,000 rpm for 5 min. The supernatant was again discarded, and the pellet was dried at 55°C until the ethanol evaporated. Finally, the pellet was resuspended in 50 µL of Tris-EDTA buffer and incubated at 37°C.

Nested PCR

Nested PCR was employed to assess the presence of deletions in flies, utilizing genomic DNA as the template. 1 µL of DNA was used for the initial PCR with an outer set of primers (Key resources table). Following this, 1 µL of the initial PCR product was employed for a subsequent PCR with an inner set of primers. The resulting PCR products were then resolved on 0.8% agarose gel to visualize the desired bands indicative of deletions.

Measurement of developmental timing and pupariation

Flies were permitted to lay eggs for 3 hours on agar plates supplemented with yeast, synchronizing the larvae. Newly hatched L1 larvae were then collected and transferred to vials. Pupariation times and dates were recorded daily during the light cycle. Data from 8-10 vials were aggregated, organized by pupariation time and cumulative percentage pupariation, and analyzed using Microsoft Excel.

Antibody generation and western blotting

To generate antibodies against Drosophila Nup107 (CG6743), the N-terminal 210 amino acids, recognized as the most unique and antigenic region, were sub-cloned into the modified tag-less pET28a(+) vector, pET28a(+)-JK vector. The protein was expressed in Escherichia coli BL21 (DE3) cells, induced with 200 μM IPTG (Sigma), and incubated at 30°C for 4 hours. Following cell pelleting, the pellet was resuspended in lysis buffer (50 mM Tris pH 8.0, 1 mM EDTA, and 25% Sucrose) with 1× protease inhibitor mixture (Roche Applied Science), and lysozyme (from 50 mg/mL stock) was added to facilitate bacterial lysis. After sonication and centrifugation, the pellet was successively resuspended in inclusion body buffer I (20 mM Tris pH 8.0, 0.2 M NaCl, 1% Sodium deoxycholate, and 2 mM EGTA) and buffer II (10 mM Tris pH 8.0, 0.25% Sodium deoxycholate and 1 mM EGTA) and centrifuged. This process was repeated three times within inclusion body buffer II, and the final pellet was dissolved in 8 M Urea buffer (10 mM Tris-HCl pH 8.0, 8 M Urea, 0.1 mM NaN3, and 1 mM EGTA), diluted to 6 M Urea, and centrifuged. The supernatant was loaded onto SDS-PAGE, and the desired band was cut for protein elution. Rabbit polyclonal antibodies against Nup107 were generated at Bio Bharati Life Sciences Pvt. Ltd., Kolkata, West Bengal, India. The antibodies obtained were subjected to affinity purification. The purification involved the chemical cross-linking of purified antigens to N-hydroxy succinimidyl-Sepharose beads (Sigma). The elution process was carried out under low pH conditions, followed by neutralization. Subsequently, the eluted antibodies were dialyzed against phosphate-buffered saline (PBS) overnight at 4°C to remove any remaining impurities and optimize their stability and functionality for subsequent experimental use.

Larval brain complexes from third-instar larvae were dissected, lysed in Laemmli buffer, and resolved on 8% SDS-PAGE. Two equivalent head complexes were loaded per well. After transfer to a polyvinylidene difluoride membrane, blocking was performed with 5% fat-free milk. The membrane was then incubated with polyclonal anti-Nup107 antibody (1:500) and anti-α-tubulin (DSHB, 12G10) (1:5000) at 4°C overnight. Following three washes with TBS-T buffer (20 mM Tris-HCl pH 7.5, 150 mM NaCl, 0.1% Tween-20), the membrane was incubated with secondary antibodies, anti-rabbit-Alexa-Fluor Plus 680 (Thermo Fisher Scientific #A32734) and anti-mouse-Alexa-Fluor Plus 800 (Thermo Fisher Scientific #A32730). Following incubation with secondary antibodies, the membrane was washed three times for 10 minutes each with Tris-buffered saline containing Tween-20 (TBS-T) to remove unbound antibodies. The washed membrane was then subjected to imaging using the Li-Cor IR system (Model: 9120), allowing for the visualization and analysis of the protein bands on the membrane.

Immunostaining

For immunostaining Drosophila salivary glands, the previously reported protocol was followed (Mehta et al., 2021). The larvae were dissected in cold phosphate-buffered saline (PBS) to isolate salivary glands. Glands underwent pre-extraction with 1% PBST (PBS + 1% Triton X-100), then fixation in freshly prepared 4% paraformaldehyde for 30 minutes at room temperature. Subsequently, the glands were thoroughly washed with 0.3% PBST (0.3% Triton X-100 containing PBS). Blocking was performed for 1 h with 5% normal goat serum (#005-000-001, The Jackson Laboratory, USA). The glands were stained with the following primary antibodies: anti-Nup107 (1:100), mAb414 (1:500, Bio-legend) and anti-EcR (1:20, DDA2.7, DSHB) overnight at 4°C. Tissues were washed three times with 0.3% PBST (PBS + 0.3% Triton X-100), followed by incubation with secondary antibodies: anti-rabbit Alexa Fluor 488 (1:800, #A11034, Thermo Fisher Scientific), anti-rabbit Alexa Fluor 568 (1:800, #A11036, Thermo Fisher Scientific), anti-mouse Alexa Fluor 488 (1:800, #A11029, Thermo Fisher Scientific), anti-mouse Alexa Fluor 568 (1:800, #A11004, Thermo Fisher Scientific), and anti-rabbit Alexa Fluor 647 (1:800, Jackson ImmunoResearch). Following secondary antibody incubation, tissues were washed three times with 0.3% PBST (PBS + 0.2% Triton X-100) and mounted with DAPI-containing Fluoroshield (#F6057, Sigma). The same protocol was followed for staining of the Brain complex and prothoracic gland.

Fluorescence Intensity quantification

Images were acquired using an Olympus Confocal Laser Scanning Microscope FV3000. Subsequent image processing was conducted using the Fiji software, and signal intensities were averaged using GraphPad software (Prism). To quantify the EcR nuclear to cytoplasmic ratio, three distinct regions of interest (ROI) were designated per nucleus and its surrounding cytoplasm. Fiji software measured the average intensity in each ROI, and the mean of these intensities was determined per nucleus and its adjacent cytoplasm. The final graph depicts the ratio of the mean intensities observed in the nucleus to that in the cytoplasm. All experiments were conducted with a minimum of three independent replicates.

Quantitative RT-PCR

Total RNA was extracted from various genotypes (control, ubiquitously depleted Nup107, and prothoracic gland-specific depletion of Nup107) of whole late third instar larvae utilizing the total tissue RNA isolation kit (Favorgen Biotech). One microgram of total RNA was employed for cDNA synthesis with the iScript cDNA synthesis kit (Bio-Rad). The resulting cDNA was diluted fivefold, and 1 μL from each genotype was utilized as a template. RT-PCR was conducted using SYBR Green PCR master mix on an Applied Biosystems QuantStudio 3 Real-Time PCR System. The Rpl69 gene served as the control, and relative transcript levels were determined using the CT value (2-ΔΔCT). Differences in gene expression were analyzed using Student’s t-test, with a p-value < 0.05 considered significant. The primers used are detailed in the Key resources table. The resultant graph illustrated the fold change in gene expression, plotted using GraphPad software (Prism).

20-hydroxyecdysone (20E) level measurements

Three biological replicates of 25 mg larvae from each genotype were collected at the specified times after egg laying (AEL). The larvae were washed, dried, and weighed before being flash-frozen on dry ice and stored at −80°C. Ecdysone extraction was done by thoroughly homogenizing the frozen samples in 300 µL ice-cold methanol with a plastic pestle. After centrifugation at 17,000 x g for 10 minutes, the supernatant was divided into two Eppendorf tubes containing approximately 150 µL supernatant. Methanol from both tubes was evaporated in a vacuum centrifuge for 60 minutes, and the resulting pellets were redissolved by adding 200 µL Enzyme Immunoassay (EIA) buffer to one of the tubes. After vortexing, the same 200 µL EIA buffer was transferred to the second tube, followed by another round of vortexing. The Enzyme-Linked Immunosorbent Assay (ELISA) was performed using a 20-hydroxyecdysone ELISA kit (#EIAHYD) from Thermo Fisher Scientific.

20E Rescue Experiment

20-hydroxyecdysone (#H5142, Sigma) was dissolved in ethanol to achieve a 5 mg/mL concentration. Salivary glands from third instar larvae of different genotypes were extracted and then incubated for 6 hours in Schneider’s S2 media with 50 μM 20E. After the incubation, the glands underwent procedures for immunostaining and RNA extraction.

To facilitate the rescue of the larvae through feeding, 30-third instar larvae, Nup107 depleted larvae cultured at 29°C incubator, were first washed with water. They were then transferred to vials containing either 20E (at a final concentration of 0.2 mg/mL) or 95% ethanol (in the same amount as the 20E). Once the larvae were added to the vials, these were returned to the 29°C incubator and observed for pupariation, recording the time of this developmental stage.

eLife Assessment

This useful study presents findings on the developmental roles of Nup107, a key nucleoporin, in regulating the larval-to-pupal transition in Drosophila melanogaster through its involvement in ecdysone signaling. The evidence supporting the authors' claims is solid, with robust experimental approaches including RNAi knockdown and rescue experiments. The findings highlight Nup107's function in regulating ecdysone biosynthesis, specifically through the regulation of EcR levels and Halloween genes expression in the prothoracic gland; additionally, rescue experiments suggest that the RTK PTTH/Torso signaling pathway is disrupted upon Nup107 depletion, further emphasizing its role in ecdysone regulation. However, finding a mechanism, addressing potential off-target effects of RNAi, and exploring alternative mutant models would strengthen the findings as the currently proposed mechanism is not fully supported by the data.

Reviewer #1 (Public review):

This study provides a thorough analysis of Nup107's role in Drosophila metamorphosis, demonstrating that its depletion leads to developmental arrest at the third larval instar stage due to disruptions in ecdysone biosynthesis and EcR signaling. Importantly, the authors establish a novel connection between Nup107 and Torso receptor expression, linking it to the hormonal cascade regulating pupariation.

However, some contradictory results weaken the conclusions of the study. The authors claim that Nup107 is involved in the translocation of EcR from the cytoplasm to the nucleus. However, the evidence provided in the paper suggests it more likely regulates EcR expression positively, as EcR is undetectable in Nup107-depleted animals, even below background levels. Additionally, the link between Nup107 and Torso is not fully substantiated. While overexpression of Torso appears to rescue the lack of 20E production in the prothoracic gland, the distinct phenotypes of Torso and Nup107 depletion-developmental delay in the former versus complete larval arrest in the latter complicate understanding of Nup107's precise role.

To clarify these discrepancies, further investigation into whether Nup107 interacts with other critical signaling pathways related to the regulation of ecdysone biosynthesis, such as EGFR or TGF-β, would be beneficial and could strengthen the findings.

In summary, although the study presents some intriguing observations, several conclusions are not well-supported by the experimental data.

Reviewer #2 (Public review):

Summary:

The manuscript by Kawadkar et al investigates the role of Nup107 in developmental progression via the regulation of ecdysone signaling. The authors identify an interesting phenotype of Nup107 whole-body RNAi depletion in Drosophila development - developmental arrest at the late larval stage. Nup107-depleted larvae exhibit mislocalization of the Ecdysone receptor (EcR) from the nucleus to the cytoplasm and reduced expression of EcR target genes in salivary glands, indicative of compromised ecdysone signaling. This mis-localization of EcR in salivary glands was phenocopied when Nup107 was depleted only in the prothoracic gland (PG), suggesting that it is not nuclear transport of EcR but the presence of ecdysone (normally secreted from PG) that is affected. Consistently, whole-body levels of ecdysone were shown to be reduced in Nup107 KD, particularly at the late third instar stage when a spike in ecdysone normally occurs. Importantly, the authors could rescue the developmental arrest and EcR mislocalization phenotypes of Nup107 KD by adding exogenous ecdysone, supporting the notion that Nup107 depletion disrupts biosynthesis of ecdysone, which arrests normal development. Additionally, they found that rescue of the Nup107 KD phenotype can also be achieved by over-expression of the receptor tyrosine kinase torso, which is thought to be the upstream regulator of ecdysone synthesis in the PG. Transcript levels of the torso are also shown to be downregulated in the Nup107KD, as are transcript levels of multiple ecdysone biosynthesis genes. Together, these experiments reveal a new role of Nup107 or nuclear pore levels in hormone-driven developmental progression, likely via regulation of levels of torso and torso-stimulated ecdysone biosynthesis.

Strengths:

The developmental phenotypes of an NPC component presented in the manuscript are striking and novel, and the data appears to be of high quality. The rescue experiments are particularly significant, providing strong evidence that Nup107 functions upstream of torso and ecdysone levels in the regulation of developmental timing and progression.

Weaknesses:

The underlying mechanism is however not clear, and any insight into how Nup107 may regulate these pathways would greatly strengthen the manuscript. Some suggestions to address this are detailed below.

Major questions:

(1) Determining how specific this phenotype is to Nup107 vs. to reduced NPC levels overall would give some mechanistic insight. Does knocking down other components of the Nup107 subcomplex (the Y-complex) lead to similar phenotypes? Given the published gene regulatory function of Nup107, do other gene regulatory Nups such as Nup98 or Nup153 produce these phenotypes?

(2) In a related issue, does this level of Nup107 KD produce lower NPC levels? It is expected to, but actual quantification of nuclear pores in Nup107-depleted tissues should be added. These and the above experiments would help address a key mechanistic question - is this phenotype the result of lower numbers of nuclear pores or specifically of Nup107?

(3) Additional experiments on how Nup107 regulates the torso would provide further insight. Does Nup107 regulate transcription of the torso or perhaps its mRNA export? Looking at nascent levels of the torso transcript and the localization of its mRNA can help answer this question. Or alternatively, does Nup107 physically bind the torso?

(4) The depletion level of Nup107 RNAi specifically in the salivary gland vs. the prothoracic gland should be compared by RT-qPCR or western blotting.

(5) The UAS-torso rescue experiment should also include the control of an additional UAS construct - so Nup107; UAS-control vs Nup107; UAS-torso should be compared in the context of rescue to make sure the Gal4 driver is functioning at similar levels in the rescue experiment.

Minor:

(6) Figures and figure legends can stand to be more explicit and detailed, respectively.

Reviewer #3 (Public review):

Summary:

In this study by Kawadkar et al, the authors investigate the developmental role of Nup107, a nucleoporin, in regulating the larval-to-pupal transition in Drosophila through RNAi knockdown and CRISPR-Cas9-mediated gene editing. They demonstrate that Nup107, an essential component of the nuclear pore complex (NPC), is crucial for regulating ecdysone signaling during developmental transitions. The authors show that the depletion of Nup107 disrupts these processes, offering valuable insights into its role in development.

Specifically, they find that:

(1) Nup107 depletion impairs pupariation during the larval-to-pupal transition.

(2) RNAi knockdown of Nup107 results in defects in EcR nuclear translocation, a key regulator of ecdysone signaling.

(3) Exogenous 20-hydroxyecdysone (20E) rescues pupariation blocks, but rescued pupae fail to close.

(4) Nup107 RNAi-induced defects can be rescued by activation of the MAP kinase pathway.

Strengths:

The manuscript provides strong evidence that Nup107, a component of the nuclear pore complex (NPC), plays a crucial role in regulating the larval-to-pupal transition in Drosophila, particularly in ecdysone signaling.

The authors employ a combination of RNAi knockdown, CRISPR-Cas9 gene editing, and rescue experiments, offering a comprehensive approach to studying Nup107's developmental function.

The study effectively connects Nup107 to ecdysone signaling, a key regulator of developmental transitions, offering novel insights into the molecular mechanisms controlling metamorphosis.

The use of exogenous 20-hydroxyecdysone (20E) and activation of the MAP kinase pathway provides a strong mechanistic perspective, suggesting that Nup107 may influence EcR signaling and ecdysone biosynthesis.

Weaknesses:

The authors do not sufficiently address the potential off-target effects of RNAi, which could impact the validity of their findings. Alternative approaches, such as heterozygous or clonal studies, could help confirm the specificity of the observed phenotypes.

NPC Complex Specificity: While the authors focus on Nup107, it remains unclear whether the observed defects are specific to this nucleoporin or if other NPC components also contribute to similar defects. Demonstrating similar results with other NPC components would strengthen their claims.

Although the authors show that Nup107 depletion disrupts EcR signaling, the precise molecular mechanism by which Nup107 influences this process is not fully explored. Further investigation into how Nup107 regulates EcR nuclear translocation or ecdysone biosynthesis would improve the clarity of the findings.

There are some typographical errors and overly strong phrases, such as "unequivocally demonstrate," which could be softened. Additionally, the presentation of redundant data in different tissues could be streamlined to enhance clarity and flow.

Author response:

The following is the authors’ response to the original reviews

Public Reviews:

Reviewer #1 (Public review):

This study provides a thorough analysis of Nup107's role in Drosophila metamorphosis, demonstrating that its depletion leads to developmental arrest at the third larval instar stage due to disruptions in ecdysone biosynthesis and EcR signaling. Importantly, the authors establish a novel connection between Nup107 and Torso receptor expression, linking it to the hormonal cascade regulating pupariation.

However, some contradictory results weaken the conclusions of the study. The authors claim that Nup107 is involved in the translocation of EcR from the cytoplasm to the nucleus. However, the evidence provided in the paper suggests it more likely regulates EcR expression positively, as EcR is undetectable in Nup107-depleted animals, even below background levels.

We appreciate the concern raised in this public review. However, we must clarify that we do not claim that Nup107 directly regulates the translocation of EcR from the cytoplasm to nucleus, rather Nup107 regulates Ecdysone hormone (20E) synthesis which in turn affects EcR translocation. In the manuscript, we posited this hypothesis if Nup107 will regulate EcR nuclear translocation (9th line of 2nd paragraph on page 6). We have spelled this out more clearly as the 3rd subsection title of the Results section, and in the discussion (8th line of 2nd paragraph on page 11).

20E acts through the EcR to induce the transcription of EcR responsive genes including the EcR. This creates a positive autoregulatory loop that enhances the EcR level through ecdysone signaling (1). Since Nup107 depletion leads to a reduction in ecdysone levels, it disrupts the transcription autoregulatory EcR expression loop. This can contribute to the reduced EcR levels seen in Nup107-depleted animals.

Additionally, the link between Nup107 and Torso is not fully substantiated. While overexpression of Torso appears to rescue the lack of 20E production in the prothoracic gland, the distinct phenotypes of Torso and Nup107 depletion-developmental delay in the former versus complete larval arrest in the latter complicate understanding of Nup107's precise role.

We understand that there are differences in the developmental delay when Tosro and Nup107 depletion is analyzed. However, the two molecules being compared here are very different, and variability in their depletion could contribute observed phenotypic differences (2). Even if there is no variability of depletion of Torso and Nup107­­­, we believe that Nup107, being more widely expressed, and involved in the regulation of various cellular processes, induces stronger defects.

Further, we think that RNAi-mediated depletion of Nup107 in prothoracic glands (PG) causes significant reduction in the PG size, which may exert a pronounced defect in 20E biosynthesis through the Halloween genes, inducing a stronger developmental arrest.

To clarify these discrepancies, further investigation into whether Nup107 interacts with other critical signaling pathways related to the regulation of ecdysone biosynthesis, such as EGFR or TGF-β, would be beneficial and could strengthen the findings.

In summary, although the study presents some intriguing observations, several conclusions are not well-supported by the experimental data.

We agree with the reviewer’s suggestion. As noted in the literature, five RTKs-torso, InR, EGFR, Alk, and Pvr-stimulate the PI3K/Akt pathway, which plays a crucial role in the PG functioning and controlling pupariation and body size (3). We have checked the torso and EGFR signaling. We rescued Nup107 defects with the torso overexpression, however, constitutively active EGFR (BL-59843) did not rescue the phenotype (data was not shown). Nonetheless, we plan to examine the EGFR pathway activation by measuring the pERK levels in Nup107-depleted PGs.

Reviewer #2 (Public review):

Summary:

The manuscript by Kawadkar et al investigates the role of Nup107 in developmental progression via the regulation of ecdysone signaling. The authors identify an interesting phenotype of Nup107 whole-body RNAi depletion in Drosophila development - developmental arrest at the late larval stage. Nup107-depleted larvae exhibit mis-localization of the Ecdysone receptor (EcR) from the nucleus to the cytoplasm and reduced expression of EcR target genes in salivary glands, indicative of compromised ecdysone signaling. This mis-localization of EcR in salivary glands was phenocopied when Nup107 was depleted only in the prothoracic gland (PG), suggesting that it is not nuclear transport of EcR but the presence of ecdysone (normally secreted from PG) that is affected. Consistently, whole-body levels of ecdysone were shown to be reduced in Nup107 KD, particularly at the late third instar stage when a spike in ecdysone normally occurs. Importantly, the authors could rescue the developmental arrest and EcR mislocalization phenotypes of Nup107 KD by adding exogenous ecdysone, supporting the notion that Nup107 depletion disrupts biosynthesis of ecdysone, which arrests normal development. Additionally, they found that rescue of the Nup107 KD phenotype can also be achieved by over-expression of the receptor tyrosine kinase torso, which is thought to be the upstream regulator of ecdysone synthesis in the PG. Transcript levels of the torso are also shown to be downregulated in the Nup107KD, as are transcript levels of multiple ecdysone biosynthesis genes. Together, these experiments reveal a new role of Nup107 or nuclear pore levels in hormone-driven developmental progression, likely via regulation of levels of torso and torso-stimulated ecdysone biosynthesis.

Strengths:

The developmental phenotypes of an NPC component presented in the manuscript are striking and novel, and the data appears to be of high quality. The rescue experiments are particularly significant, providing strong evidence that Nup107 functions upstream of torso and ecdysone levels in the regulation of developmental timing and progression.

Weaknesses:

The underlying mechanism is however not clear, and any insight into how Nup107 may regulate these pathways would greatly strengthen the manuscript. Some suggestions to address this are detailed below.

Major questions:

(1) Determining how specific this phenotype is to Nup107 vs. to reduced NPC levels overall would give some mechanistic insight. Does knocking down other components of the Nup107 subcomplex (the Y-complex) lead to similar phenotypes? Given the published gene regulatory function of Nup107, do other gene regulatory Nups such as Nup98 or Nup153 produce these phenotypes?

We thank this public review for raising this concern. Working with a Nup-complex like the Nup107 complex, this concern is anticipated but difficult to address as many Nups function beyond their complex identity. Our observations with all other members of the Nup107-complex, including dELYS, suggest that except Nup107, none of the other tested Nup107-complex members could induce larval developmental arrest.

In this study, we primarily focused on the Nup107 complex (outer ring complex) of the NPC. However, previous studies have reported that Nup98 and Nup153 interact with chromatin, with these investigations conducted in Drosophila S2 cells (4, 5, 6). We have now examined other nucleoporins outside of this complex, such as Nup153.

We ubiquitously depleted Nup153 using the Actin5C-Gal4 driver and assessed the pupariation profile of the knockdown larvae in comparison to control larvae. In contrast to the Nup107 knockdown, when Nup153 is depleted to less than 50% levels, no impact on pupariation was observed (Auhtor response image 1)

Nup153 depletion does not affect the Drosophila metamorphosis.

(2) In a related issue, does this level of Nup107 KD produce lower NPC levels? It is expected to, but actual quantification of nuclear pores in Nup107-depleted tissues should be added. These and the above experiments would help address a key mechanistic question - is this phenotype the result of lower numbers of nuclear pores or specifically of Nup107?

We agree with the concern raised here, and to address the concern raised here, we stained the control and Nup107 depleted salivary glands with mAb414 antibody (exclusively FG-repeat Nup recognizing antibody). While Nup107 intensities are significantly reduced at the nuclear envelope in Nup107 depleted salivary glands, the mAb414 staining seems unperturbed (Author response image 2).

Nup107 depletion does not perturb overall NPC composition.

Comparison of salivary gland nucleus upon control and Nup107 knockdown. The Nup107 is shown in green and mAb414, staining for other FG-repeat containing nucleoporins is shown in red. Scale bars, 5µm.

(3) Additional experiments on how Nup107 regulates the torso would provide further insight. Does Nup107 regulate transcription of the torso or perhaps its mRNA export? Looking at nascent levels of the torso transcript and the localization of its mRNA can help answer this question. Or alternatively, does Nup107 physically bind the torso?

While the concern regarding torso transcript level is genuine, we have already reported in the manuscript that Nup107 directly regulates torso expression. When Nup107 is depleted, torso levels go down, which in turn controls ecdysone production and subsequent EcR signaling (Figure 6B of the manuscript).

However, the exact nature of Nup107 regulation on torso expression is still unclear. Since the Nup107 is known to interact with chromatin (7), it may affect torso transcription. The possibility of a stable and physiologically relevant interaction between Nup107 and the torso in a cellular context is unlikely largely due to their distinct subcellular localizations. If we investigate this further, it will require a significant amount of time for having reagents and experimentation, and currently stands beyond the scope of this manuscript.

(4) The depletion level of Nup107 RNAi specifically in the salivary gland vs. the prothoracic gland should be compared by RT-qPCR or western blotting.

Although we know that the Nup107 protein signal is reduced in SG upon knockdown (Figure 3B), we have not compared the Nup107 transcript level in these two tissues (SG and PG) upon RNAi. As suggested here, we evaluated the knockdown efficiency of Nup107 using the salivary gland-specific driver AB1-Gal4 and the prothoracic gland-specific driver Phm-Gal4. Our results indicate a significant reduction in Nup107 transcript levels upon Nup107 RNAi in both SG and PG compared to their respective controls (Author response image 3).

Nup107 levels are significantly reduced upon Nup107KK RNAi.

Quantification of Nup107 transcript levels from control and Nup107 depleted larvae [tissue specific depletion using AB1-Gal4 (A) and Phm-Gal4 (B)]. Data are represented from at least three independent experiments. Statistical significance was derived from the Student’s t-test. Error bars represent SEM. **p = <0.004

(5) The UAS-torso rescue experiment should also include the control of an additional UAS construct - so Nup107; UAS-control vs Nup107; UAS-torso should be compared in the context of rescue to make sure the Gal4 driver is functioning at similar levels in the rescue experiment.

This is a very valid point, and we took this into account while planning the experiment. In such cases, often the GAL4 dilution can be critical. We have demonstrated in Figure S7, that GAL4 dilution is not blurring our observations. We used the Nup107KK; UAS-GFP as control alongside the Nup107KK; UAS-torso. We conclude that the presence of GFP signals in prothoracic glands and their reduced size indicates genes downstream to both UAS sequences are transcribed, and GAL4 dilution does not play a role here.

Minor:

(6) Figures and figure legends can stand to be more explicit and detailed, respectively.

We have revisited all figures and their corresponding legends to ensure appropriate and explicit details are provided.

Reviewer #3 (Public review):

Summary:

In this study by Kawadkar et al, the authors investigate the developmental role of Nup107, a nucleoporin, in regulating the larval-to-pupal transition in Drosophila through RNAi knockdown and CRISPR-Cas9-mediated gene editing. They demonstrate that Nup107, an essential component of the nuclear pore complex (NPC), is crucial for regulating ecdysone signaling during developmental transitions. The authors show that the depletion of Nup107 disrupts these processes, offering valuable insights into its role in development.

Specifically, they find that:

(1) Nup107 depletion impairs pupariation during the larval-to-pupal transition.

(2) RNAi knockdown of Nup107 results in defects in EcR nuclear translocation, a key regulator of ecdysone signaling.

(3) Exogenous 20-hydroxyecdysone (20E) rescues pupariation blocks, but rescued pupae fail to close.

(4) Nup107 RNAi-induced defects can be rescued by activation of the MAP kinase pathway.

Strengths:

The manuscript provides strong evidence that Nup107, a component of the nuclear pore complex (NPC), plays a crucial role in regulating the larval-to-pupal transition in Drosophila, particularly in ecdysone signaling.

The authors employ a combination of RNAi knockdown, CRISPR-Cas9 gene editing, and rescue experiments, offering a comprehensive approach to studying Nup107's developmental function.

The study effectively connects Nup107 to ecdysone signaling, a key regulator of developmental transitions, offering novel insights into the molecular mechanisms controlling metamorphosis.

The use of exogenous 20-hydroxyecdysone (20E) and activation of the MAP kinase pathway provides a strong mechanistic perspective, suggesting that Nup107 may influence EcR signaling and ecdysone biosynthesis.

Weaknesses:

The authors do not sufficiently address the potential off-target effects of RNAi, which could impact the validity of their findings. Alternative approaches, such as heterozygous or clonal studies, could help confirm the specificity of the observed phenotypes.

This is a very valid point raised, and we are aware of the consequences of the off-target effects of RNAi. To assert the effects of authentic RNAi and reduce the off-target effects, we have used two RNAi lines (Nup107GD and Nup107KK) against Nup107. Both RNAi induced comparable levels of Nup107 reduction, and using these lines, ubiquitous and PG specific knockdown produced similar phenotypes. Although the Nup107GD line exhibited a relatively stronger knockdown compared to the Nup107KK line, we preferentially used the Nup107KK line because the Nup107GD line is based on the P-element insertion, and the exact landing site is unknown. Furthermore, there is an off-target predicted for the Nup107GD line, where a 19bp sequence aligns with the bifocal (bif) sequence. The bif-encoded protein is involved in axon guidance and regulation of axon extension. However, the Nup107KK line does not have a predicted off-target molecule, and we know its precise landing site on the second chromosome. Thus, the Nup107KK line was ultimately used in experimentation for its clearer and more reliable genetic background.

We are also investigating Nup107 knockdown in the prothoracic gland, which exhibits polyteny. Additionally, the number of cells in the prothoracic gland is quite limited, approximately 50-60 cells (8). Given this, there is a possibility that a clonal study may not yield the phenotype.

NPC Complex Specificity: While the authors focus on Nup107, it remains unclear whether the observed defects are specific to this nucleoporin or if other NPC components also contribute to similar defects. Demonstrating similar results with other NPC components would strengthen their claims.

We thank this public review for raising this concern. Working with a Nup-complex like the Nup107 complex, this concern is anticipated but difficult to address as many Nups function beyond their complex identity. Our observations with all other members of the Nup107-complex, including dELYS, suggest that except Nup107, none of the other Nup107-complex members could induce larval developmental arrest. Since the study is primarily focused on the Nup107 complex (outer ring complex) of the NPC, we have not examined many more nucleoporins outside of this complex. But our observations with Nup153 knockdown, a nuclear basket nucleoporin, is comparable to control, with no delay in development (Author response image 1)

Although the authors show that Nup107 depletion disrupts EcR signaling, the precise molecular mechanism by which Nup107 influences this process is not fully explored. Further investigation into how Nup107 regulates EcR nuclear translocation or ecdysone biosynthesis would improve the clarity of the findings.

We appreciate the concern raised. Through our observation, we have proposed the upstream effect of Nup107 on the PTTH-torso-20E-EcR axis regulating developmental transitions. We know that Nup107 regulates torso levels, but we do not know if Nup107 directly interacts with torso. We would like to address whether Nup107 exerts control on PTTH levels also.

However, we must emphasize that Nup107 does not directly regulate the translocation of EcR. On the contrary, we have demonstrated that when Nup107 is depleted only in the salivary gland, EcR translocates into the nucleus. Thus we conclude that the EcR translocation is 20E dependent and Nup107 independent. Further, we have argued that Nup107 regulates the expression of Halloween genes required for ecdysone biosynthesis. We are interested in identifying if Nup107 associates directly or through some protein to chromatin to bring about the changes in gene expression required for normal development.

There are some typographical errors and overly strong phrases, such as "unequivocally demonstrate," which could be softened. Additionally, the presentation of redundant data in different tissues could be streamlined to enhance clarity and flow.

Response: We thank the reviewer for this observation. We have put our best efforts to remove all typographical errors and have now made more reasonable statements based on our conclusions.

Recommendations for the Authors:

Reviewer #1 (Recommendations for the authors):

The manuscript presents compelling evidence that Nup107 plays a role in regulating ecdysone production. However, significant concerns remain regarding the effects on EcR localization and expression, as well as the claimed link between PTTH/Torso signaling and Nup107's function, as the evidence provided is not conclusive.

The hypothesis that Nup107 mediates EcR translocation from the cytoplasm to the nucleus appears misinterpreted by the authors. Based on the presented images, particularly for the prothoracic gland (PG) Figure 3C, Nup107 depletion seems to impact EcR protein levels rather than its localization. This conclusion is supported by data showing that EcR transcripts are autonomously downregulated in the absence of Nup107. Furthermore, the restoration of nuclear EcR levels upon exogenous 20E supplementation suggests that (1) Nup107 is dispensable for EcR activation and function, and (2) its primary role lies in regulating ecdysone production.

We appreciate the concern raised by reviewer. However, we must clarify that we do not claim that Nup107 directly regulates the translocation of EcR from the cytoplasm, rather Nup107 regulates Ecdysone hormone (20E) synthesis which in turn affects EcR translocation. In the manuscript, we posited this hypothesis if Nup107 will regulate EcR nuclear translocation (9th line of 2nd paragraph on page 6). We have spelled this out more clearly as the 3rd subsection title of the Results section, and in the discussion (8th line of 2nd paragraph on page 11).

20E acts through the EcR to induce the transcription of EcR responsive genes including the EcR. This creates a positive autoregulatory loop that enhances the EcR level through ecdysone signaling (1). Since Nup107 depletion leads to a reduction in ecdysone levels, it disrupts the transcription autoregulatory EcR expression loop. This can contribute to the reduced EcR levels seen in Nup107-depleted animals.

Given that nucleoporins are known to influence mRNA transport-for instance, Nup107 has been shown to control Scn5a mRNA transport (Guan et al., 2019)-the observed effects on Halloween gene and EcR expression may stem from disruptions in mRNA transport to the cytoplasm. The downregulation of Shade further supports this hypothesis, as restricted ecdysone biosynthesis typically induces Shade upregulation in peripheral tissues. Quantifying potential mRNA accumulation in the nuclei of PG cells in Nup107-depleted animals would clarify this.

The reviewer raised a valid point, and we fully agree with the concern that Nup107 has been shown to control Scn5a mRNA transport (Guan et al., 2019). The observed effects on Halloween gene and EcR expression could indeed stem from disruptions in efficient mRNA export to the cytoplasm. However, if Nup107 were regulating the mRNA export of Halloween genes and EcR, we should not expect a rescue of the Nup107 developmental delay phenotype with torso overexpression. But, by overexpressing the torso in the Nup107 depletion background, we are activating the torso pathway dependent Halloween gene expression, and rescuing the developmental delay phenotype of Nup107 depletion.

With the current data, it is difficult to conclusively claim a role for Nup107 in EcR translocation or expression. Additional experiments, such as EcR overexpression in Nup107-depleted animals or Nup107 overexpression, would help determine its precise role.

We appreciate the concern raised by reviewer. We did attempt to rescue the Nup107 depletion phenotype by overexpressing EcR (BL-6868) in the Nup107-RNAi background. However, we were unable to rescue the Nup107 depletion dependent developmental delay phenotype with this approach. This further suggests that the phenotype is not merely due to low level of EcR, but it is due to low availability of ecdysone hormone and EcR signaling.

The second major issue is the proposed link between Nup107 and PTTH/Torso signaling. The authors suggest that Nup107 regulates ecdysone production through Torso expression based on rescue experiments. However, this is inconsistent with the distinct phenotypes observed when Nup107 or Torso signaling is disrupted. While PTTH/Torso signaling causes only a modest developmental delay (12 hours to 2 days, depending on the mutant), Nup107 depletion results in a complete developmental arrest at the larval stage. This discrepancy raises doubts about the assertion that Torso overexpression alone rescues such a severe phenotype. One possibility is that PTTH levels are upregulated in Nup107-depleted animals, leading to overactivation of the pathway when Torso is overexpressed. Quantifying PTTH levels in Nup107-depleted animals could address this.

The reviewer raised a valid point, and we fully acknowledge this concern. While we do not completely agree with the idea of PTTH upregulation in Nup107 depleted larvae, as suggested here, we believe that quantifying PTTH levels upon Nup107 depletion can provide a useful insight. To address it, we quantified PTTH levels in Nup107-depleted larvae and found no significant change in PTTH expression compared to controls (Author response image 4).

Nup107 knockdown does not affect the PTTH level.

Quantitation of PTTH transcript levels from control and Nup107 depleted larvae (Prothoracic specific depletion Phm-Gal4). Data are represented from at least three independent experiments. Statistical significance was derived from the Student's t-test. ns is non-significant.

Another possibility is that the stock used for Torso overexpression, which includes a trk mutant, may introduce genetic interactions that overactivate the pathway. Using a clean UAS-Torso stock would resolve this issue.

We appreciate the reviewer’s observation regarding the use of the Torso overexpression line (BL-92604), which carries the trk null allele on the second chromosome. The cleaved form of the trk serves as ligand for the troso receptor. Since it may serve as ligand for the torso, I am not sure how trk null allele bearing line when used along for torso overexpression studies will overactivate the pathway.

We realized this concern and the fly line used in this study and reported in the manuscript was generated through the following genetic strategy using the BL-92604 line. First, a double balancer stock (Sco/CyO; MKRS/TM6.Tb) was used to generate the Sco/CyO; UAS-torso/ UAS-torso genotype. This recombinant line was subsequently combined with the Nup107KK line. Through the use of the double balancer strategy, we effectively replaced Nup107 RNAi genotype on the second chromosome, thereby ensuring that our final experimental setup is free from trk mutant contamination, if at all.

Moreover, the rescue of Nup107 depletion phenotypes by RasV12 overexpression suggests that multiple RTKs, not just Torso, are affected. EGFR signaling, the primary regulator of ecdysone biosynthesis in the PG during the last larval stage, is notably absent from the authors' analysis. EGFR inactivation is known to arrest development, and previous studies indicate that Nup107 can reduce EGFR pathway activity (Kim et al, 2010). The authors should analyze EGFR pathway activity in the absence of Nup107. Overexpressing EGF ligands like Vein or Spitz in the PG (rather than the receptor) in a Nup107-depleted background would provide more relevant insights.

The RasGTPase is one of the common effector molecules downstream of an activated receptor kinase. Rescue with a constitutively activated form of RasGTPase (RasV12) suggests one of the routes which is activated downstream of the torso receptor. It does not directly suggest all different RTKs are affected and are involved. Our idea of performing a rescue experiment was to see if the pathway activated downstream of the torso involves RasGTPase.

As noted in the literature, five RTKs—torso, InR, EGFR, Alk, and Pvr—stimulate the PI3K/Akt pathway, which plays a crucial role in the PG for controlling pupariation and body size (3). Although EGFR signaling is important, PTTH/Torso signaling is considered the primary mediator of metamorphic timing. In response to the suggestion to analyze EGFR pathway activity in the absence of Nup107, we attempted to rescue the phenotype by overexpressing constitutively active EGFR (BL-59843) in the Nup107-depleted background (data was not shown). We used constitutively active EGFR to bypass the availability of its ligands (vein and spitz). Unfortunately, we were unable to rescue the phenotype with this approach, which further suggests that EGFR is not the targeted RTK pathway in this context. By rescuing with torso, we found that Nup107 regulates torso-mediated Ras/Erk signaling to control metamorphosis.

Additional issues require clarification:

(1) RNAi Efficiency: In Figure 1C, the Nup107GD line shows a stronger knockdown effect than Nup107KK, yet most experiments were conducted with the weaker line. This might explain the residual Nup107 protein observed in Figure 2. Could the authors justify this choice?

This is a very valid point raised, and we are aware of the consequences of the off-target effects of RNAi. To assert the effects of authentic RNAi and reduce the off-target effects, we have used two RNAi lines (Nup107GD and Nup107KK) against Nup107. Both RNAi induced comparable levels of Nup107 reduction, and using these lines, ubiquitous and PG specific knockdown produced similar phenotypes. Although the Nup107GD line exhibited a relatively stronger knockdown compared to the Nup107KK line, we preferentially used the Nup107KK line because the Nup107GD line is based on the P-element insertion, and the exact landing site is unknown. Furthermore, there is an off-target predicted for the Nup107GD line, where a 19bp sequence aligns with the bifocal (bif) sequence. The bif-encoded protein is involved in axon guidance and regulation of axon extension. However, the Nup107KK line does not have a predicted off-target molecule, and we know its precise landing site on the second chromosome. Thus, the Nup107KK line was ultimately used in experimentation for its clearer and more reliable genetic background.

(2) Control Comparisons: In Figure 3, the effects of Nup107 depletion on EcR expression in salivary glands (SG) and PG are shown, but only SG controls are provided. Including PG controls would enable proper comparisons. These controls should also be added to Figures 5, 6, and S5.

As suggested by the reviewer, we have checked the EcR localization in prothoracic gland (Author response image 5), also. As shown in figure R5, when PGs isolated from control, Nup107-RNAi and torso overexpression in Nup107 background were stained for EcR, the observations made were indistinguishable from those made in SGs of the indicated genetic combinations. This indicated that Nup107 regulates EcR signaling by regulating the 20E biosynthesis.

Prothoracic gland’s specific torso expression rescues EcR nuclear translocation defects.

Immunofluorescence-based detection of nucleocytoplasmic distribution of EcR (EcR antibody, red) in control, prothoracic gland specific Nup107 knockdown (Phm-Gal4>Nup107KK) and torso overexpressing PG-specific Nup107 knockdown (Phm-Gal4>Nup107KK; UAS-torso) third instar larval Prothoracic gland nuclei. DNA is stained with DAPI. Scale bars, 20 μm.

(3) Clarify the function of Torso in the text: The authors must revise their description of Torso signaling as the primary regulator of ecdysone production in both the results and discussion sections. Specifically, in the results section, the claim that Torso depletion induces developmental arrest is inaccurate. Instead, available evidence, including Rewitz et al. 2009, demonstrates that Torso depletion causes a delay of approximately five days rather than a complete developmental arrest. This discrepancy should be corrected to avoid overstating the role of Torso signaling in ecdysone regulation and to align the manuscript with established findings.

We agree with the reviewer. We have incorporated the suggestion at the relevant place in the main manuscript.

Reviewer #3 (Recommendations for the authors):

These findings suggest that Nup107 is involved in regulating ecdysone signaling during developmental transitions, with depletion of Nup107 disrupting hormone-regulated processes. Moreover, the rescue experiments hint that Nup107 might directly influence EcR signaling and ecdysone biosynthesis, though the precise molecular mechanism remains unclear.

Overall, the manuscript presents compelling data supporting Nup107's role in regulating developmental transitions. However, I have a few comments for consideration:

Major Comments:

RNAi Specificity: While RNAi is a powerful tool, the authors do not sufficiently address potential off-target effects, which could undermine the conclusions. Although a mutant Nup107 is described, it is lethal-are heterozygous or clonal studies possible to validate the findings more robustly?

This is a very valid point raised, and we are aware of the consequences of the off-target effects of RNAi. To assert the effects of authentic RNAi and reduce the off-target effects, we have used two RNAi lines (Nup107GD and Nup107KK) against Nup107. Both RNAi induced comparable levels of Nup107 reduction, and using these lines, ubiquitous and PG specific knockdown produced similar phenotypes. Although the Nup107GD line exhibited a relatively stronger knockdown compared to the Nup107KK line, we preferentially used the Nup107KK line because the Nup107GD line is based on the P-element insertion, and the exact landing site is unknown. Furthermore, there is an off-target predicted for the Nup107GD line, where a 19bp sequence aligns with the bifocal (bif) sequence. The bif-encoded protein is involved in axon guidance and regulation of axon extension. However, the Nup107KK line does not have a predicted off-target molecule, and we know its precise landing site on the second chromosome. Thus, the Nup107KK line was ultimately used in experimentation for its clearer and more reliable genetic background.

Following the suggestion from the reviewer, we considered conducting heterozygous and clonal analyses using the Nup107 mutant. We have carried out Nup107 knockdown studies in the prothoracic gland, which has a limited number of cells (50-60 cells) and is known to exhibit polyteny (8). Keeping these aspects of the Prothoracic gland in mind, the possibility that a clonal study will yield the phenotype is scarce. However, we will consider moving forward with this approach also.

(2) NPC Complex Specificity: It remains unclear whether the observed defects are specific to Nup107 or if other NPC components also cause similar defects. If the authors are unable to use Nup107 mutants, they could demonstrate similar defects with other critical NPC members to bolster their claim.

We thank this public review for raising this concern. Working with a Nup-complex like the Nup107 complex, this concern is anticipated but difficult to address as many Nups function beyond their complex identity. Our analysis of Nup153 depleted organisms indicates no developmental delay/defect. We have also assessed effects of knockdown of all other members of the Nup107-complex, including dELYS, but except Nup107 no other member of the Nup107-complex could induce developmental arrest in the third instar stage causing lack of pupariation. However, the null mutant of Nup133, the direct interactor of Nup107 in the Nup107-complex, induces a delay in pupariation (unpublished data).

(3) Molecular Mechanism of EcR Signaling: The manuscript shows that Nup107 depletion affects EcR signaling and ecdysone biosynthesis, but the molecular basis of this regulation is not fully explored. Does phosphorylated ERK (p-ERK) fail to enter the nucleus? Clarifying this mechanism would strengthen the study's impact.

We appreciate the reviewer’s insightful comment and fully agree with the concern. To address this, we examined the subcellular localization of phosphorylated ERK (p-ERK) in the prothoracic gland of control larvae, Nup107-depleted larvae, and Nup107-depleted larvae with torso overexpression. In control larvae, p-ERK was predominantly localized in the nucleus. However, in Nup107-depleted larvae, p-ERK was largely retained in the cytoplasm, indicating impaired pathway activation and nuclear translocation. Notably, overexpression of the torso in the Nup107-depleted background restored nuclear localization of p-ERK in the prothoracic gland (Author response image 6). These findings suggest that Nup107 regulates Drosophila metamorphosis, in part, through modulation of torso-mediated MAPK signaling.

Nup107 regulates torso activation dependent p-ERK localization.

Detection of nucleocytoplasmic distribution of p-ERK (anti- p-ERK antibody, green) in the third instar larval prothoracic glands of control, PG-specific Nup107 knockdown (Phm-Gal4>Nup107KK) and PG-specific torso overexpression in Nup107 knockdown background (Phm-Gal4>Nup107KK; UAS-torso). DNA is stained with DAPI. Scale bars, 20 µm.

Minor Comments:

(1) The manuscript contains typographical errors that may hinder readability. Additionally, some phrases (e.g., "unequivocally demonstrate") may be overly strong. Consider adjusting language to reflect the nature of the data more accurately.

We agree with the reviewer. We have edited the manuscript accordingly to crease out such typographical errors at relevant places in the main manuscript.

(2) The data presentation could be improved by eliminating redundancy. Some sections repeat similar findings in different tissues, which could be consolidated to improve clarity and flow.

While we agree with the comment, we could not help ourselves in tissue redundancy for presenting our data for EcR translocation studies. I wish we could use another tissue. However, we have put EcR localization and p-ERK translocation data in the responses to present another non-redundant tissue perspective (Figures R5 and R6).

References:

(1) Varghese, Jishy, and Stephen M Cohen. “microRNA miR-14 acts to modulate a positive autoregulatory loop controlling steroid hormone signaling in Drosophila.” Genes & development vol. 21,18 (2007): 2277-82. doi:10.1101/gad.439807

(2) Rewitz, Kim F et al. “The insect neuropeptide PTTH activates receptor tyrosine kinase torso to initiate metamorphosis.” Science (New York, N.Y.) vol. 326,5958 (2009): 1403-5. doi:10.1126/science.1176450

(3) Pan, Xueyang, and Michael B O'Connor. “Coordination among multiple receptor tyrosine kinase signals controls Drosophila developmental timing and body size.” Cell reports vol. 36,9 (2021): 109644. doi:10.1016/j.celrep.2021.109644

(4) Pascual-Garcia, Pau et al. “Metazoan Nuclear Pores Provide a Scaffold for Poised Genes and Mediate Induced Enhancer-Promoter Contacts.” Molecular cell vol. 66,1 (2017): 63-76.e6. doi:10.1016/j.molcel.2017.02.020

(5) Pascual-Garcia, Pau et al. “Nup98-dependent transcriptional memory is established independently of transcription.” eLife vol. 11 e63404. 15 Mar. 2022, doi:10.7554/eLife.63404

(6) Kadota, Shinichi et al. “Nucleoporin 153 links nuclear pore complex to chromatin architecture by mediating CTCF and cohesin binding.” Nature communications vol. 11,1 2606. 25 May. 2020, doi:10.1038/s41467-020-16394-3

(7) Gozalo, Alejandro et al. “Core Components of the Nuclear Pore Bind Distinct States of Chromatin and Contribute to Polycomb Repression.” Molecular cell vol. 77,1 (2020): 67-81.e7. doi:10.1016/j.molcel.2019.10.017

(8) Shimell, MaryJane, and Michael B O'Connor. “Endoreplication in the Drosophila melanogaster prothoracic gland is dispensable for the critical weight checkpoint.” microPublication biology vol. 2023 10.17912/micropub.biology.000741. 21 Feb. 2023, doi:10.17912/micropub.biology.000741

Version 2

Abstract

Abstract

Nuclear pore complexes (NPCs), composed of nucleoporins (Nups), affect nucleocytoplasmic transport, thus influencing cell division and gene regulation. Nup107 subcomplex members have been studied in housekeeping functions, diseases, and developmental disorders.

We report a unique regulatory function for Nup107 in metamorphic transition during Drosophila development. RNAi-mediated Nup107-depleted larvae were arrested in the third-instar larval stage with no signs of pupariation. This lack of pupariation is primarily due to inhibited nuclear translocation and transcriptional activation by EcR. We demonstrate the involvement of Nup107 in the transcription of the Halloween genes, modulating ecdysone biosynthesis and the EcR pathway activation. The regulation of EcR-mediated metamorphosis by the receptor tyrosine kinase, torso, is well documented. Accordingly, overexpression of the torso and MAP-kinase pathway activator, rasV12, in the Nup107 depletion background rescues the phenotypes, implying that Nup107 is an epistatic regulator of Torso-mediated activation of EcR signaling during metamorphosis.

Introduction

The multi-protein assembly of nucleoporins (Nups) constitutes nuclear pore complexes (NPCs). In eukaryotic cells, NPCs serve as molecular conduits for the trafficking of proteins and RNAs between the nucleus and the cytoplasm. The NPC composition, as ascertained from proteomic analyses conducted in yeasts and vertebrates, has contributed to our understanding, revealing that approximately 30 distinct Nups are arranged in sub-structures and are distributed on different faces of NPCs (Cronshaw et al., 2002; Rout et al., 2000). The substructures are the outer cytoplasmic and inner nuclear rings, which surround a central inner ring called the scaffold ring, primarily aiding in the process of nucleocytoplasmic transport (Lin & Hoelz, 2019).

The largest Nup107 sub-complex, also called the Y-complex, is symmetrically located on both sides of the nuclear membrane. In metazoans, the Y-complex is composed of ten distinct Nups (ELYS, Nup160, Nup133, Nup107, Nup96, Nup85, Nup43, Nup37, Sec13, and Seh1). ELYS is a sub-stoichiometric Y-complex member and is present only on the nucleoplasmic side (D’Angelo & Hetzer, 2008; Morchoisne-Bolhy et al., 2015). Nup107, along with Nup133, forms the stalk of the Y-complex and is critically required for the Y-complex stability. The Nup107 complex plays a pivotal role in facilitating ELYS-coordinated post-mitotic NPC assembly (Boehmer et al., 2003; Walther et al., 2003) and messenger RNA (mRNA) export (Bai et al., 2004). Additionally, the Nup107 complex members actively participate in mitosis, contributing to the regulation of kinetochore-microtubule polymerization (Mishra et al., 2010; Zuccolo et al., 2007). The multitude of cellular processes in which the Y-complex performs vital functions suggests it to be a central component in maintaining cellular homeostasis.

As a stable constituent of NPC, many Nups, including the members of the Y-complex, associate with chromatin and exert transcriptional regulation. Notably, interactions between active genes and Nups occurring predominantly within the nucleoplasm have been reported for dynamic Nups such as Nup98, ELYS, and Sec13 (Capelson, Liang, et al., 2010; Kalverda et al., 2010; Kuhn et al., 2019). In Drosophila, the dual nucleoporin, ELYS, governs the development, and ELYS RNAi-induced developmental defects are due to the reactivation of the Dorsal (NF-κB) pathway even during the late larval stages (Mehta et al., 2020).

Nup107 is associated with actively transcribing genes at the nuclear periphery (Gozalo et al., 2020). The disruption of Nup107 in zebrafish embryos leads to significant developmental anomalies, including the absence of the pharyngeal skeleton (Zheng et al., 2012). Moreover, the biallelic Nup107 mutations (D157Y and D831A) correlate well with clinical conditions such as microcephaly and steroid-resistant nephrotic Syndrome (Miyake et al., 2015). In Drosophila, Nup107 co-localizes with Lamin during meiotic division, and Nup107 depletion perturbs Lamin localization, leading to a higher frequency of cytokinesis failure during male meiosis (Hayashi et al., 2016). Nup107 influences the regulation of cell fate in aged and transformed cells by modulating EGFR signaling and the nuclear trafficking of ERK protein (Kim et al., 2010).

Drosophila undergoes an elaborate metamorphosis initiated by the neuropeptide prothoracicotropic hormone (PTTH) (Rewitz et al., 2013). Bilateral neurons projecting into the prothoracic gland (PG), when stimulated by the PTTH, induce Ecdysone production, which is subsequently released into the circulatory system for conversion by peripheral tissues into its active form, 20-hydroxyecdysone (20E) (Johnson et al., 2013; McBrayer et al., 2007; Shimell et al., 2018). The 20E binds to the ecdysone receptor, and the whole complex translocates into the nucleus and binds to chromatin to activate ecdysone-inducible genes (Johnston et al., 2011; Kozlova & Thummel, 2002; Tennessen & Thummel, 2011). In the prothoracic gland, the primary neuroendocrine organ, PTTH signals through the receptor tyrosine kinase (RTK), Torso. The Torso-dependent activation of the MAP-kinase pathway is responsible for the production and release of the ecdysone hormone. However, ecdysone synthesis can also be regulated by the EGFR pathway (Cruz et al., 2020; Yamanaka et al., 2013). While Nup107 modulates EGFR pathway activation (Kim et al., 2010) but the involvement of EGFR and torso pathways in ecdysone-dependent metamorphosis is undeniable. We dissect the involvement of Nup107 in Torso-mediated signaling and underlying mechanisms during Drosophila metamorphosis.

In a reverse genetic RNAi screening for Nup107 complex members, we noted that Nup107 RNAi induces a significant developmental arrest at the third instar larval stage. Further analysis revealed that the Ecdysone Receptor (EcR) signaling pathway is perturbed, and EcR fails to translocate into the nucleus in Nup107 knockdown. The failure of the EcR nuclear localization upon Nup107 depletion is due to significantly reduced ecdysone hormone levels during the late third instar larval stage. Interestingly, overexpression of the torso and the rasV12 in Nup107-depleted larvae rescued the developmental arrest and subsequently initiated pupariation. We propose that Nup107, an epistatic regulator of torso-pathway activation in the PG, enables ecdysone surge for efficient metamorphic transition.

Results

Nup107 is essential for the larval to pupal metamorphic transition

Cell biological analyses in the mammalian cell culture system have shed light on the critical regulatory roles of Nup107 in vertebrates. Yet its importance in development is poorly understood. In this context, we started the characterization of Drosophila Y-complex member Nup107 in greater detail. Utilizing RNAi lines (Nup107KK and Nup107GD), we performed ubiquitous depletion through Actin5C-GAL4. Interestingly, Nup107 depletion, inducing the third instar larval arrest, results in the complete cessation of pupariation (120 h AEL, right panel, Figure 1A), which was accompanied by an extension of larval feeding and growth periods. Quantitative assessment of Nup107 transcript levels in the Nup107GD and Nup107KK RNAi lines suggested efficient Nup107 knockdown (approximately 60-70%, Figure 1B). For further analyses, we generated polyclonal antibodies against the Nup107 amino-terminal antigenic fragment (amino acids 1–210; see methods for details). Purified anti-Nup107 polyclonal antibodies detected a band of approximately ∼100 kDa in lysates prepared from the control larval brain complex. The intensity of this ∼100 kDa band was significantly reduced in lysates prepared from organisms where Nup107 was knocked down ubiquitously (Figure 1C) using Actin5C-GAL4 driving Nup107 RNAi (denoted as ubiquitous hereon). Further, the immunostaining with Nup107 antibodies identified a conserved and robust nuclear rim staining pattern overlapping with mAb414 antibodies recognizing FG-nucleoporins (Figure S1A) and mRFP-tagged Nup107 expressed through its endogenous promoter (Figure S1B) in salivary gland tissues. These observations confirm the efficacy of Nup107 knockdown and provide a handle to assess levels and localization of Nup107 in affected tissues. To further investigate the role of Nup107 in development, we generated a Nup107 null mutant using CRISPR-Cas9-mediated gene editing. This comprehensive approach involving RNAi-mediated knockdown and CRISPR-Cas9 gene editing is expected to provide valuable insights into the significance of Nup107 in Drosophila development. The gRNAs targeting regions close to the start and stop codons of the nup107 gene generated knockout (Nup107KO) mutants, which were confirmed by sequencing and PCR (Figures S2A, and S2B). However, the Nup107KO mutants could not be used as the Nup107KO induces lethality at the embryonic stage. So, we carried out all the analyses hereon with Nup107 RNAi lines.

Nup107 depletion impairs metamorphosis.

Analysis of Nup107 depletion, and its impact on growth and development of organism.

(A) Growth profile of third instar larvae from Actin5C-Gal4 driven control and Nup107 knockdowns (Nup107GD and Nup107KK RNAi lines) at 96 h AEL (after egg laying) and 120 h AEL.

(B) Quantitation of Nup107 knockdown efficiency. Data are represented from at least three independent experiments. Statistical significance was derived from the Student’s t-test. Error bars represent SEM. **p = <0.001 and ****p = <0.0001.

(C) Immunodetection of Nup107 protein levels in third instar larval brain-complex lysates from control and Nup107 knockdown.

(D) Quantification of Nup107 protein levels seen in (C). Data are represented from at least three independent experiments. Statistical significance was derived from the Student’s t-test. Error bars represent SEM. ***p = <0.0002 and ****p = <0.0001.

(E) Comparison of pupariation profiles of control and Nup107 knockdown organisms.

Nup107 contributes significantly to ecdysone signaling

The depletion of Nup107 in Drosophila resulted in a distinct halt in growth and a developmental arrest at the third instar stage (Figure 1A). To discern the pathways affecting juvenile to adult developmental transition in Drosophila, we focused on levels of the sole insect steroid hormone, ecdysone.

We examined the localization of EcR in the late third instar salivary glands of both the control and Nup107-depleted larvae (using the Nup107KK line). Wild-type larvae exhibited normal EcR localization within the nucleus, but the nuclear translocation of EcR is perturbed in the Nup107-depleted larvae, with the bulk of the signal retained in the cytoplasm (Figures 2A, and 2B). Quantitative analysis of EcR intensities in the cytoplasm and nucleus further established a significant decrease in EcR signals inside the nucleus upon Nup107 depletion (Figure 2C). This observation may suggest that Nup107 is required for EcR nuclear localization to mediate the critical larval to pupal developmental transition. Furthermore, we noticed that significantly smaller-sized salivary glands and the brain complex in Nup107-depleted larvae (Figure S3).

Ubiquitous knockdown of Nup107 disrupts ecdysone signaling.

Assessment of ecdysone receptor dependent signaling in salivary glands of third instar larvae.

(A-B) Staining of third instar larval salivary glands from control (A) and ubiquitous Nup107 knockdown (B) with EcR (anti-EcR antibody, red) and Nup107 (anti-Nup107 antibody, green). DNA is stained with DAPI. Scale bars, 20 μm. Charts represent the line scan intensity profile of EcR (Red) and DAPI (Cyan) in the salivary gland nucleus region.

(C) Quantification of the nucleocytoplasmic ratio of EcR under control and Nup107 knockdown conditions. At least 45 nuclei were analyzed from 7 to 8 pairs of salivary glands. Statistical significance was derived from the Student’s t-test. Error bars represent SEM. ****p = < 0.0001.

(D-F) Analysis of ecdysone-inducible genes, EcR (D), Eip75A (E), and Eip74EF (F) expression, respectively, at the onset of metamorphosis (late third instar larvae stage). Data are represented from at least three independent experiments. Statistical significance was derived from the Student’s t-test. The error bars represent the SEM.

***p = <0.0004 and ****p = <0.0001.

Prompted by the cytoplasmic accumulation of EcR, we investigated whether the mRNA levels of ecdysone-inducible genes were affected. Under normal conditions, when 20E binds with EcR, the activated EcR occupies the ecdysone response element (ERE) in the promoter region of genes. The ecdysone receptor is thought to function in a positive auto-regulatory loop, which may elevate EcR levels and sustain ecdysone signaling (Varghese & Cohen, 2007). Firstly, we examined the EcR transcript levels, revealing a reduction under Nup107-depletion conditions (Figure 2D). Subsequently, we measured the mRNA levels of known EcR target genes, Eip75A and Eip74EF, to find that the expression of each of these two target genes is reduced and correlates with Nup107 knockdown levels (Figures 2E, and 2F). Similar results were observed in the depletion of Nup107 using the GD-based RNAi line (Figure S4). The defects observed during metamorphic transitions can be attributed to impaired ecdysone signaling in Nup107-depleted organisms, prompting a closer examination of their regulatory relationship.

Nup107 is dispensable for the nuclear import of EcR in the target tissue

We established that the nuclear localization of EcR is impaired upon Nup107 knockdown (see Figure 2B). Typically, EcR nuclear localization follows an intricate mechanism where the 20E binding allows nuclear translocation of EcR and subsequent activation of its target genes (Cronauer et al., 2007; Johnston et al., 2011; Lenaerts et al., 2019). Biosynthesis of ecdysone hormone from dietary cholesterol is a prerequisite for EcR activation, and requires a group of P450 enzymes coded by the Halloween genes (Kannangara et al., 2021; Niwa & Niwa, 2014). Moreover, the involvement of nucleoporin, Nup358, in facilitating nuclear transport of Met juvenile hormone receptors is well documented (He et al., 2017). Consequently, we posited the hypothesis that Nup107 either regulates active 20E-EcR complex formation by affecting 20E biosynthesis in ubiquitous Nup107 knockdown scenarios or directly regulates EcR nuclear translocation in the target tissue.

We chose salivary glands to address these hypotheses and depleted Nup107 using salivary gland-specific AB1-GAL4 and PG-specific Phm-GAL4. Surprisingly, in contrast to the ubiquitous knockdown of Nup107, nuclear localization of EcR remained unaffected in salivary gland-specific Nup107 knockdown (Figures 3A, 3B, S5A, and S5B). Interestingly, the PG-specific Nup107 knockdown phenocopied ubiquitous Nup107 knockdown-dependent EcR nuclear localization defects (Figures 3C, and S5C). Quantification of EcR signals from Nup107-depleted late third instar salivary gland cells and comparison with control salivary glands indicates that the nuclear/cytoplasmic ratios are drastically reduced in the case of PG-specific depletion but unaltered in salivary gland-specific Nup107 depletion (Figures 3D, and S5D). Consistent with these observations, expression levels of ecdysone-inducible genes Eip75A and Eip74EF were significantly reduced in PG-specific Nup107 knockdown (Figures 3E, 3F, S5E, and S5F). In accordance with unperturbed EcR nuclear translocation, salivary gland-specific depletion of Nup107 yielded no discernible differences in larval growth and pupariation compared to the control (Figures S6A, and S6C). However, the PG-specific knockdown induced an extended third instar stage lifespan (10-12 days after egg laying, Figures S6B, and S6C). The observed decrease in ecdysone-inducible gene expression during late third-instar developmental stages can explain the potential impairment of metamorphosis induction seen upon ubiquitous or PG-specific Nup107 knockdown. In addition to the reduced size of the salivary gland and brain complex, we also noticed a compromise in the size of the Prothoracic gland upon Nup107 knockdown (Figure S7).

Targeted knockdown of Nup107 in the PG perturbs ecdysone signaling pathway.

Analyzing impact of tissue specific Nup107 depletion on ecdysone receptor signaling in third instar larval salivary glands.

(A-C) Detection of and quantitation of nucleocytoplasmic distribution of EcR (anti-EcR antibody, red) and Nup107 (anti-Nup107 antibody, green) in control (A), salivary gland-specific Nup107 depletion (B), and prothoracic gland-specific Nup107 depletion (C) from third instar larval salivary gland nuclei. DNA is stained with DAPI. Scale bars, 20 μm. Charts represent the line scan intensity profile of EcR (Red) and DAPI (Cyan) in the salivary gland nucleus region.

(C) EcR nucleo-cytoplasmic quantification ratio from the salivary gland and prothoracic gland-specific Nup107 knockdown, respectively. At least 45 nuclei were analyzed from 7 to 8 pairs of salivary glands. Statistical significance was derived from the Student’s t-test. Error bars represent SEM. ****p = <0.0001, and ns is non-significant.

(E-F) Quantitation of expression of Eip75A (E) and Eip74EF (F) ecdysone-inducible genes at the onset of metamorphosis (RNA isolated from late third instar larvae of control and prothoracic gland-specific Nup107 depletion). Data are represented from at least three independent experiments. Statistical significance was derived from the Student’s t-test. Error bars represent SEM. **p = <0.008 and ****p = <0.0001.

These observations suggest that Nup107 exerts a regulation on ecdysone biosynthesis and active 20E-EcR complex formation rather than playing a direct role in EcR nuclear translocation.

Nup107 exerts control on the EcR pathway through Ecdysone level regulation

We reasoned that Nup107 may regulate the ecdysone biosynthesis in PG to induce larval stage growth arrest. We delved into analyzing the effect of Nup107 knockdown on ecdysone production. The considerable decrease in PG size due to Nup107 knockdown (Figure S7) can potentially reduce 20E production. This prompted us to explore the potential role of Nup107 in influencing ecdysone production, and we assessed the impact of Nup107 knockdown on ecdysone biosynthesis.

Utilizing an ELISA-based detection method, we assessed 20E levels in larvae at 96 and 120 hours after egg laying (AEL). Larvae from different experimental conditions, including control, ubiquitous Nup107-depletion, and PG-specific Nup107-depletion, were used in this analysis. Strikingly, the results indicated a substantial decrease in total 20E levels upon Nup107 knockdown (approximately 3-fold and 10-fold, respectively, for ubiquitous and PG-specific knockdown), particularly at 120 AEL, which coincides with the 20E surge seen during metamorphosis (Figures 4A and S8A). This observation is crucial and suggests a potential defect in ecdysone biosynthesis in Nup107-depleted organisms.

Nup107 critically regulates the expression of ecdysone-biosynthetic genes.

(A) ELISA measurements of whole-body 20-hydroxyecdysone (20E) levels in control, ubiquitous (Actin5C-Gal4), and prothoracic gland-specific (Phm-Gal4) Nup107 depletion at 96 and 120 h AEL. Data are represented from at least three independent experiments. Statistical significance was derived from the Student’s t-test. Error bars represent SEM. **p = <0.0013, ***p = < 0.0009 and ns is non-significant.

(B) Schematic representation of a Prothoracic gland cell showing genes involved in ecdysone biosynthesis from cholesterol.

(C-G) Quantification of ecdysone-biosynthetic gene expression levels of Spookier (C), Phantom (D), Disembodied (E), Shadow (F), and Shade (G) in cDNA isolated from control, and Nup107 knockdown late third instar larvae. Data are represented from at least three independent experiments. Statistical significance was derived from the Student’s t-test. Error bars represent SEM. **p = <0.001, ***p = <0.0005 and, ****p = <0.0001.

As depicted in Figure 4B, the PTTH hormone signaling in the PG upregulates the expression of Halloween genes (spookier, phantom, disembodied, and shadow) responsible for ecdysone biosynthesis, and the shade gene product is required for active 20E generation in peripheral tissues (Christensen et al., 2020; McBrayer et al., 2007; Shimell et al., 2018). We analyzed the Halloween genes transcript level in Nup107-depleted larvae and observed a significant downregulation for each of the Halloween quartet genes mentioned earlier. (Figures 4C-4F, and S8B-S8E). Further, the level of the shade gene was also reduced twofold in Nup107-depleted larvae (Figures 4G, and S8F).

20E rescues Nup107-dependent EcR localization defects

The observed correlation between the expression levels of ecdysone biosynthetic genes and reduced levels of 20E upon Nup107 knockdown strongly suggests a role for Nup107 in 20E biosynthesis, active 20E-EcR complex formation, and subsequent nuclear translocation. It is not surprising in this context that exogenous ecdysone supplementation through larval food rescues pupariation blocks arising from ecdysone deficiency (Garen et al., 1977; Ou et al., 2016; Shimell et al., 2018). We experimented the same under Nup107-depletion conditions and supplemented exogenous ecdysone to PG-specific Nup107-depleted larvae by feeding them with a diet enriched with 20-hydroxyecdysone (0.2 mg/ml). Supplementation of 20E to Nup107-depleted larvae significantly alleviated the developmental arrest, and the onset of pupariation was comparable to the control (Table S1). However, none of the pupae could eventually eclose successfully, probably due to other effects of Nup107 depletion. We asked if nuclear translocation of EcR can also be rescued by exogenous supplementation of 20E. While incubation of salivary glands of late third instar larva in S2 media control alone did not rescue EcR localization in any of the Nup107-knockdown genotypes (Figures 5A-5C, and S9A), we noticed a complete EcR nuclear translocation rescue in ubiquitous as well as PG-specific Nup107-depleted salivary glands when incubated with 20E (Figures 5D-5F, and S9B).

20E supplementation rescues Nup107 depletion-specific EcR signaling defects.

Immunofluorescence analysis of EcR localization in 20E supplemented and non-supplemented third instar larval salivary glands.

(A-F) Visualization of the nucleocytoplasmic distribution of EcR (anti-EcR antibody, red) without 20E (A-C) and with 20E (D-F) treatment in larval salivary glands of control, ubiquitous (Actin5C-Gal4) Nup107 knockdown, and prothoracic gland-specific (Phm-Gal4) Nup107 knockdown. DNA is stained with DAPI. Scale bars, 20 μm. Charts represent the line scan intensity profile of EcR (Red) and DAPI (Cyan) in the salivary gland nucleus region.

(G-H) Comparative quantification of expression of ecdysone-inducible genes Eip75A (G) and Eip74EF (H) from 20E-treated salivary glands. Data are represented from at least three independent experiments. Statistical significance was derived from the Student’s t-test. Error bars represent SEM. *p = <0.02, **p = <0.002, ***p = <0.0008 and ns is non-significant.

We reason that the exogenous supplementation of 20E leads to active 20E-EcR complex formation as mRNA levels of the Eip75A and Eip74EF target genes were rescued significantly back to normal levels (Figures 5G, and 5H). Overall, these findings highlight the important regulatory function of Nup107 in the ecdysone signaling pathway.

Torso is an effector of Nup107-mediated functions in metamorphosis

Next, we explored the mechanism of Nup107-driven regulation of 20E levels and metamorphosis. The cell surface receptors of the tyrosine kinase family bind to ligands (growth factors and hormones) and activate signaling to regulate metabolism, cell growth, and development. Among these RTKs, the Torso, belonging to the platelet-derived growth factor receptor (PDGFR) class, plays a significant role during metamorphosis by serving as a receptor for the neuropeptide PTTH in the Drosophila brain (Sopko & Perrimon, 2013).

Functional engagement of PTTH-Torso activates the MAP kinase pathway, involving components of Ras, Raf, MEK, and extracellular signal-regulated kinase (ERK), thereby initiating metamorphosis (Figure 6A). The torso knockdown in the prothoracic glands resulted in a significant delay in the onset of pupariation, extending the developmental period by approximately 6 days, resembling the developmental arrest seen with Nup107 knockdown. Feeding 20E to torso-depleted larvae completely rescued developmental delay and normal growth phenotypes (Rewitz et al., 2009). We observed a significant decrease in the torso levels (approximately four-fold) when Nup107 was depleted ubiquitously using Actin5C-GAL4 (Figure 6B), suggesting an epistatic regulation by Nup107 on the torso. Further, the phenotypic similarity between Nup107 and torso-depletion scenarios and diminished torso level upon Nup107-depletion prompted us to investigate whether the torso is an effector of Nup107.

Torso and Nup107 act synergistically to activate the ecdysone signaling.

(A) A model of the Torso pathway and its components.

(B) Quantitation of torso transcript levels from control and Nup107 depleted larvae (ubiquitous depletion using Actin5C-Gal4). Data are represented from at least three independent experiments. Statistical significance was derived from the Student’s t-test. Error bars represent SEM. ****p = <0.0001.

(C-D) Comparison of pupariation profiles of control, Nup107 knockdown, and torso and rasV12 overexpressing rescue organisms.

(E-G) Detection and quantitation of nucleocytoplasmic distribution of EcR (anti-EcR antibody, red) and Nup107 (anti-Nup107 antibody, green) in control, torso overexpressing ubiquitous Nup107 knockdown (Actin5C-Gal4>Nup107KK; UAS-torso) and torso overexpressing PG-specific Nup107 knockdown (Phm-Gal4>Nup107KK; UAS-torso) third instar larval salivary gland nuclei. DNA is stained with DAPI. Scale bars, 20 μm. Charts represent the line scan intensity profile of EcR (Red) and DAPI (Cyan) in the salivary gland nucleus region.

(H-I) Quantification of expression of Eip75A (H) and Eip74EF (I) ecdysone-inducible genes, respectively. Data are represented from at least three independent experiments. Statistical significance was derived from the Student’s t-test. Error bars represent SEM.

*p = <0.03, **p = <0.008, ****p = <0.0001 and ns is non-significant.

Overexpression of the torso and or rasV12 has been utilized to rescue torso pathway-mediated defects (Cruz et al., 2020; Rewitz et al., 2009). We probed the possibility of Nup107 phenotype rescue by ubiquitous or PG-specific overexpression of the torso and rasV12. Overexpression of either torso or rasV12 in Nup107-depletion background completely rescued the pupariation defects (Figures 6C, and 6D). Interestingly, the overexpression of Egfr (another RTK linked to Drosophila metamorphosis) and Usp (co-receptor with EcR) in Nup107 depletion background could not rescue the pupariation defects (data not shown), indicating that the torso overexpression-mediated rescue of Nup107-depletion phenotypes is specific. The torso overexpression-dependent rescue prompted us to analyze the status of EcR nuclear translocation, ecdysone biosynthesis, and ecdysone-inducible gene expression. The restoration of EcR nuclear translocation (Figures 6E-6G), the ecdysone biosynthesis genes, spookier, phantom, disembodied, and shadow (Figures S10B-S10E), and ecdysone target genes Eip75A and Eip74EF (Figures 6H, and 6I) to control levels indicated that torso could efficiently rescue Nup107 phenotypes.

These observations suggest that the torso is a downstream effector of Nup107 functions, and torso-dependent signaling, responsible for 20E synthesis in PG and metamorphosis, is regulated by Nup107 levels.

Discussion

Apart from its importance in mRNA export and cell division at the cellular level, the NPC’s Y-complex member Nup107, when mutated, correlates with developmental abnormalities such as microcephaly and nephrotic syndrome (Miyake et al., 2015; Zheng et al., 2012). In several human cancer types, Nup107 exhibits high expression and serves as a biomarker for hepatocellular carcinoma. A strong structural and functional conservation between human and Drosophila Nup107 proteins can help us model human diseases and gain mechanistic insights into Nup107 functions in Drosophila (Shore et al., 2022; Weinberg-Shukron et al., 2015). We observe that Nup107 is involved in critical developmental transitions from larva to pupa during Drosophila development as pupariation is completely arrested (Figure 1). The foraging third instar larva must acquire a critical weight and sufficient energy stores before it can metamorphose into a non-feeding pupa, molting into a healthy adult. The activation of the Torso receptor and subsequent surge in ecdysone synthesis are essential for larval to pupal transition (Christensen et al., 2020; Hao et al., 2021; Luo et al., 2024). The elevated ecdysone levels trigger the translocation of the heterodimeric ecdysone receptor [in complex with Ultraspiracle (Usp)] into the nucleus to induce pupariation-specific transcriptional programming (Johnston et al., 2011). Similarly, when Nup107 is depleted, nuclear translocation of EcR is abolished, arresting metamorphosis at the level of the third instar larval stage (Figure 2).

Translocation of nuclear receptor proteins like the ecdysone receptor to the nucleus is a crucial step for transcriptional activation. The marked absence of EcR from the nucleus was intriguing, as Nup107 generally participates in the nuclear export process. Perhaps Nup107 regulates pupariation by modulating ecdysone synthesis, ecdysone signaling, and nuclear translocation of EcR. Surprisingly, Nup107 is dispensable for the nuclear translocation of EcR in the target tissue, the salivary glands (Figure 3). Nup107 exerts strong control over the expression of Halloween genes involved in ecdysone biosynthesis, resulting in diminished 20E titer, poor EcR activation, and delayed larva-to-pupa transition (Figures 4 and S6). The effect of Nup107 on Halloween genes adheres well with pupariation defects. These observations are in concurrence with reports of low ecdysone levels disrupting the pupariation process, leading to a halt in insect development (Christensen et al., 2020; Cruz et al., 2020).

In addition to ecdysone signaling, insect metamorphosis also has a strong contribution from receptor tyrosine kinase RTK pathway signaling. Notably, Torso signaling is essential for embryonic development and metamorphosis in Drosophila. The receptor expression level impacts the signaling output: high receptor levels trigger a robust and transient signal, while lower levels result in a weaker, sustained signal (Jenni et al., 2015; Konogami et al., 2016). Accordingly, the reduced ecdysone levels in the torso knockdown prothoracic glands induced pupariation delays (Rewitz et al., 2009). We observe a similar defect with Nup107 knockdown, indicating a possible reduction in torso levels. Interestingly, we observed significantly reduced torso levels in Nup107-depleted organisms, suggesting an essential upstream function for Nup107 in regulating RTK pathway activation and the associated Drosophila metamorphosis process. In the contrasting observation, Nup107-depleted larval tissues (salivary glands, brain complex, and prothoracic gland) are significantly smaller in size (Figures S3, and S7). It is important to know how Nup107 contributes to organ size maintenance and if a crosstalk with RTK signaling is required in this context.

The developmental delays observed in Nup107 knockdown larvae can be restored by exogenous supplementation of 20E, suggesting that Nup107 can modulate directly or indirectly the Torso activity for ecdysone production and developmental regulation. Accordingly, we successfully rescued developmental delays by ubiquitous and PG-specific torso overexpression. Perhaps overexpression of the torso pathway mediator molecules can rescue the Nup107 developmental delay phenotypes. The oncogenic variant of the Ras-GTPase (rasV12) has been explored in a similar analysis with torso pathway mutants (Cruz et al., 2020; Rewitz et al., 2009). The alleviation of the developmental delays in the Nup107 knockdown background upon rasV12 overexpression is an indication of Nup107 serving as an epistatic regulator of the torso pathway in metamorphic transitions (Figure 6).

In essence, our findings indicate that Nup107 influences pupariation timing by regulating the torso levels, its signaling, and ecdysone biosynthesis (Figure 7). Previous research has shown that nuclear pore complexes (NPCs) are essential for maintaining global genome organization and regulating gene expression (Capelson, Doucet, et al., 2010; Capelson, Liang, et al., 2010; Iglesias et al., 2020). Particularly, Nup107 interacts with chromatin and targets active gene domains to regulate gene expression (Gozalo et al., 2020). Thus, Nup107, exerting its effects on torso transcription, is the primary regulatory event in the Drosophila metamorphosis. It is important to note that the synthesis and availability of Torso ligand, PTTH, may not be affected by the Nup107 since torso or downstream effector rasV12 overexpression is sufficient to rescue the developmental arrest phenotype. It is thus crucial to further delineate the mechanism of Nup107-dependent regulation of torso pathway activation. The whole genome transcriptomics from PG can help shed more light on the regulatory roles of Nup107. This information will offer valuable insights into how nucleoporins regulate gene expression and serve as a model for elucidating how they govern the temporal specificity of developmental processes in organisms. Furthermore, these analyses will help establish a link between Nup107, the PTTH-PG axis, and the regulation of developmental transition timing. Our observations indicate critical roles for Nup107 in both torso-ecdysone interplay in metamorphosis and torso-independent mechanisms for organ size maintenance. Together, they add valuable information to the hitherto unknown functions of the Nup107 in organismal development. Further investigations are required to identify the interactors of Nup107 involved in these coordination mechanisms in developmental transition.

Theoretical Model of Nup107 functions in metamorphosis:

During metamorphosis, the prothoracic gland (PG) responds to prothoracicotropic hormone (PTTH) via Torso receptors. The MAP kinase pathway involving Raf, MEK, and ERK, initiated by Ras, leads to Halloween gene expression responsible for Ecdysone synthesis and release. Ecdysone is converted to active 20-hydroxyecdysone (20E) in peripheral tissues. The binding of 20E to the Ecdysone receptor (EcR) allows EcR nuclear translocation and EcR pathway activation, culminating in target gene expression, facilitating the metamorphic transition. Nup107 depletion negatively impacts Torso levels and Torso pathway activation, inducing pupariation arrest, which can be rescued by autonomous activation of the Torso pathway. This figure was created using BioRender.com.

Materials and methods

Key resources table

Fly stocks and genetics

Experimental Drosophila melanogaster stocks were reared on a standard cornmeal diet (Nutri-Fly Bloomington formulation) under controlled conditions of 25°C and 60% relative humidity unless otherwise specified. Fly lines used in this study were sourced from the Bloomington Drosophila Stock Centre (BDSC) at Indiana University or from the Vienna Drosophila Resource Center (VDRC), which are listed in the Key resources table. The UAS-GFP lines were received as a gift from Dr. Varun Chaudhary (IISER Bhopal, India). Control groups were generated through crosses between the driver line and w1118 flies. For RNA interference (RNAi) experiments, crosses were maintained at 29°C to optimize GAL4 expression. The Nup107GD line, in conjunction with Actin5C-GAL4, was specifically cultivated at 23°C to generate third-instar larvae for experimental purposes.

Transgenic fly generation

In the generation of transgenic flies containing gRNA, we employed a systematic approach. The online tool available at http://targetfinder.flycrispr.neuro.brown.edu was used for the initial sgRNA design, ensuring zero predicted off-targets. Subsequently, we utilized an additional tool available at https://www.flyrnai.org/evaluateCrispr/ to evaluate and score the predicted efficiency of sgRNAs in the targeted region. The most efficient sgRNAs, demonstrating high specificity with no predicted off-targets, were selected and cloned into the pCFD4 vector. The Fly Facility services at the Centre for Cellular and Molecular Platforms, National Center for Biological Sciences (C-CAMP-NCBS), Bangalore, India, were utilized to clone and generate transgenic flies. Primer sequences employed in this study can be found in the Key resources table.

CRISPR –Cas9 mediated mutant Generation

Virgin nanos.Cas9 (BL-54591) flies were crossed with males of gRNA transgenic lines, and approximately 10 F1 progeny males were crossed with balancer flies specific for the gene of interest. A single-line cross was established using the F2 progeny to assess the deletion of nup107. Subsequently, F3 progeny were subjected to nested PCR to confirm deletions. Positive individuals were further tested for the presence of gRNA and Cas9 transgenes, and flies lacking both were propagated into stocks.

Genomic DNA Isolation

Approximately ten flies were homogenized in 250 µL of solution A, comprising 0.1 M Tris-HCl, pH 9.0, 0.1 M EDTA, and 1% SDS, supplemented with Proteinase-K. The homogenate was incubated at 70°C for 30 min. Subsequently, 35 µL of 8 M potassium acetate was added, mixed gently without vortexing, and incubated on ice for 30 min. The homogenate was then centrifuged at 13,000 rpm at 4°C for 15 min, and the supernatant was carefully collected without disturbing precipitates or the interphase. 150 µL of Isopropanol was added to the supernatant and incubated on ice for 15 min. After centrifugation at 13,000 rpm for 5 min, the supernatant was discarded, and the pellet was washed with 1 mL of 70% ethanol by centrifuging at 13,000 rpm for 5 min. The supernatant was again discarded, and the pellet was dried at 55°C until the ethanol evaporated. Finally, the pellet was resuspended in 50 µL of Tris-EDTA buffer and incubated at 37°C.

Nested PCR

Nested PCR was employed to assess the presence of deletions in flies, utilizing genomic DNA as the template. 1 µL of DNA was used for the initial PCR with an outer set of primers (Key resources table). Following this, 1 µL of the initial PCR product was employed for a subsequent PCR with an inner set of primers. The resulting PCR products were then resolved on 0.8% agarose gel to visualize the desired bands indicative of deletions.

Measurement of developmental timing and pupariation

Flies were permitted to lay eggs for 3 hours on agar plates supplemented with yeast, synchronizing the larvae. Newly hatched L1 larvae were then collected and transferred to vials. Pupariation times and dates were recorded daily during the light cycle. Data from 8-10 vials were aggregated, organized by pupariation time and cumulative percentage pupariation, and analyzed using Microsoft Excel.

Antibody generation and western blotting

To generate antibodies against Drosophila Nup107 (CG6743), the N-terminal 210 amino acids, recognized as the most unique and antigenic region, were sub-cloned into the modified tag-less pET28a(+) vector, pET28a(+)-JK vector. The protein was expressed in Escherichia coli BL21 (DE3) cells, induced with 200 μM IPTG (Sigma), and incubated at 30°C for 4 hours. Following cell pelleting, the pellet was resuspended in lysis buffer (50 mM Tris pH 8.0, 1 mM EDTA, and 25% Sucrose) with 1× protease inhibitor mixture (Roche Applied Science), and lysozyme (from 50 mg/mL stock) was added to facilitate bacterial lysis. After sonication and centrifugation, the pellet was successively resuspended in inclusion body buffer I (20 mM Tris pH 8.0, 0.2 M NaCl, 1% Sodium deoxycholate, and 2 mM EGTA) and buffer II (10 mM Tris pH 8.0, 0.25% Sodium deoxycholate, and 1 mM EGTA) and centrifuged. This process was repeated three times within inclusion body buffer II, and the final pellet was dissolved in 8 M Urea buffer (10 mM Tris-HCl pH 8.0, 8 M Urea, 0.1 mM NaN3, and 1 mM EGTA), diluted to 6 M Urea, and centrifuged. The supernatant was loaded onto SDS-PAGE, and the desired band was cut for protein elution. Rabbit polyclonal antibodies against Nup107 were generated at Bio Bharati Life Sciences Pvt. Ltd., Kolkata, West Bengal, India. The antibodies obtained were subjected to affinity purification. The purification involved the chemical cross-linking of purified antigens to N-hydroxy succinimidyl-Sepharose beads (Sigma). The elution process was carried out under low pH conditions, followed by neutralization. Subsequently, the eluted antibodies were dialyzed against phosphate-buffered saline (PBS) overnight at 4°C to remove any remaining impurities and optimize their stability and functionality for subsequent experimental use.

Larval brain complexes from third-instar larvae were dissected, lysed in Laemmli buffer, and resolved on 8% SDS-PAGE. Two equivalent head complexes were loaded per well. After transfer to a polyvinylidene difluoride membrane, blocking was performed with 5% fat-free milk. The membrane was then incubated with polyclonal anti-Nup107 antibody (1:500) and anti-α-tubulin (DSHB, 12G10) (1:5000) at 4°C overnight. Following three washes with TBS-T buffer (20 mM Tris-HCl pH 7.5, 150 mM NaCl, 0.1% Tween-20), the membrane was incubated with secondary antibodies, anti-rabbit-Alexa-Fluor Plus 680 (Thermo Fisher Scientific #A32734) and anti-mouse-Alexa-Fluor Plus 800 (Thermo Fisher Scientific #A32730). Following incubation with secondary antibodies, the membrane was washed three times for 10 minutes each with Tris-buffered saline containing Tween-20 (TBS-T) to remove unbound antibodies. The washed membrane was then subjected to imaging using the Li-Cor IR system (Model: 9120), allowing for the visualization and analysis of the protein bands on the membrane.

Immunostaining

For immunostaining Drosophila salivary glands, the previously reported protocol was followed (Mehta et al., 2021). The larvae were dissected in cold phosphate-buffered saline (PBS) to isolate salivary glands. Glands underwent pre-extraction with 1% PBST (PBS + 1% Triton X-100), then fixation in freshly prepared 4% paraformaldehyde for 30 minutes at room temperature. Subsequently, the glands were thoroughly washed with 0.3% PBST (0.3% Triton X-100 containing PBS). Blocking was performed for 1 h with 5% normal goat serum (#005-000-001, The Jackson Laboratory, USA). The glands were stained with the following primary antibodies: anti-Nup107 (1:100), mAb414 (1:500, Bio-legend) and anti-EcR (1:20, DDA2.7, DSHB) overnight at 4°C. Tissues were washed three times with 0.3% PBST (PBS + 0.3% Triton X-100), followed by incubation with secondary antibodies: anti-rabbit Alexa Fluor 488 (1:800, #A11034, Thermo Fisher Scientific), anti-rabbit Alexa Fluor 568 (1:800, #A11036, Thermo Fisher Scientific), anti-mouse Alexa Fluor 488 (1:800, #A11029, Thermo Fisher Scientific), anti-mouse Alexa Fluor 568 (1:800, #A11004, Thermo Fisher Scientific), and anti-rabbit Alexa Fluor 647 (1:800, Jackson ImmunoResearch). Following secondary antibody incubation, tissues were washed three times with 0.3% PBST (PBS + 0.2% Triton X-100) and mounted with DAPI-containing Fluoroshield (#F6057, Sigma). The same protocol was followed for staining of the Brain complex and prothoracic gland.

Fluorescence Intensity quantification

Images were acquired using an Olympus Confocal Laser Scanning Microscope FV3000. Subsequent image processing was conducted using the Fiji software, and signal intensities were averaged using GraphPad software (Prism). To quantify the EcR nuclear to cytoplasmic ratio, three distinct regions of interest (ROI) were designated per nucleus and its surrounding cytoplasm. Fiji software measured the average intensity in each ROI, and the mean of these intensities was determined per nucleus and its adjacent cytoplasm. The final graph depicts the ratio of the mean intensities observed in the nucleus to that in the cytoplasm. All experiments were conducted with a minimum of three independent replicates.

Quantitative RT-PCR

Total RNA was extracted from various genotypes (control, ubiquitously depleted Nup107, and prothoracic gland-specific depletion of Nup107) of whole late third instar larvae utilizing the total tissue RNA isolation kit (Favorgen Biotech). One microgram of total RNA was employed for cDNA synthesis with the iScript cDNA synthesis kit (Bio-Rad). The resulting cDNA was diluted fivefold, and 1 μL from each genotype was utilized as a template. RT-PCR was conducted using SYBR Green PCR master mix on an Applied Biosystems QuantStudio 3 Real-Time PCR System. The Rpl69 gene served as the control, and relative transcript levels were determined using the CT value (2-ΔΔCT). Differences in gene expression were analyzed using Student’s t-test, with a p-value < 0.05 considered significant. The primers used are detailed in the Key resources table. The resultant graph illustrated the fold change in gene expression, plotted using GraphPad software (Prism).

20-hydroxyecdysone (20E) level measurements

Three biological replicates of 25 mg larvae from each genotype were collected at the specified times after egg laying (AEL). The larvae were washed, dried, and weighed before being flash-frozen on dry ice and stored at -80°C. Ecdysone extraction was done by thoroughly homogenizing the frozen samples in 300 µL ice-cold methanol with a plastic pestle. After centrifugation at 17,000 x g for 10 minutes, the supernatant was divided into two Eppendorf tubes containing approximately 150 µL supernatant. Methanol from both tubes was evaporated in a vacuum centrifuge for 60 minutes, and the resulting pellets were redissolved by adding 200 µL Enzyme Immunoassay (EIA) buffer to one of the tubes. After vortexing, the same 200 µL EIA buffer was transferred to the second tube, followed by another round of vortexing. The Enzyme-Linked Immunosorbent Assay (ELISA) was performed using a 20-hydroxyecdysone ELISA kit (#EIAHYD) from Thermo Fisher Scientific.

20E Rescue Experiment

20-hydroxyecdysone (#H5142, Sigma) was dissolved in ethanol to achieve a 5 mg/mL concentration. Salivary glands from third instar larvae of different genotypes were extracted and then incubated for 6 hours in Schneider’s S2 media with 50 μM 20E. After the incubation, the glands underwent procedures for immunostaining and RNA extraction.

To facilitate the rescue of the larvae through feeding, thirty third instar larvae for each genotype were cultured in a 29°C incubator and were first washed with water. They were then transferred to vials containing either 20E (at a final concentration of 0.2 mg/mL) or 95% ethanol (in the same amount as the 20E). Once the larvae were added to the vials, these were returned to the 29°C incubator and observed for pupariation, recording the time of this developmental stage.